### 2018-03-01 docs/paper.tex
gutted template latex file

%%% template.tex
%%%
%%% This LaTeX source document can be used as the basis for your technical
%%% paper or abstract. Intentionally stripped of annotation, the parameters
%%% and commands should be adjusted for your particular paper - title,
%%% author, article DOI, etc.
%%% The accompanying ``template.annotated.tex'' provides copious annotation
%%% for the commands and parameters found in the source document. (The code
%%% is identical in ``template.tex'' and ``template.annotated.tex.'')

\documentclass[]{acmsiggraph}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\TOGonlineid{45678}
\TOGvolume{0}
\TOGnumber{0}
\TOGarticleDOI{0}
\TOGprojectURL{}
\TOGvideoURL{}
\TOGdataURL{}
\TOGcodeURL{}
\usepackage{color}
%\definecolor{red}{rgb}{0.9, 0.17, 0.31}
\usepackage{multirow}
\usepackage{subfig}
\usepackage{xcolor}
\usepackage{lipsum}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{glsllst} % My own package providing markup listing for glsl
\usepackage{rmlst}   % My own package providing markup listing for renderman
\usepackage{amsmath}
\usepackage{hyperref}

\lstset{
	backgroundcolor=\color[rgb]{0.95, 0.95, 0.95},
	tabsize=3,
	%rulecolor=,
	basicstyle=\footnotesize\ttfamily,
	upquote=true,
	aboveskip={1.5\baselineskip},
	columns=fixed,
	showstringspaces=false,
	extendedchars=true,
	breaklines=true,
	prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
	frame=none,
	aboveskip=15pt,
	belowskip=8pt,
	captionpos=t,
	showtabs=false,
	showspaces=false,
	showstringspaces=false,
	identifierstyle=\ttfamily,
	%keywordstyle=\color{red}\bfseries,
	%keywordstyle=[1]\bfseries\color{syntaxBlue},
	%keywordstyle=[2]\bfseries\color{syntaxRed},
	%keywordstyle=[3]\color{blue}\bfseries,
	%keywordstyle=[4]\bfseries\color{syntaxBlue},
	commentstyle=\color[rgb]{0.082,0.639,0.082},
	keywordstyle=[1]\bfseries\color[rgb]{0,0,0.75},
	keywordstyle=[2]\bfseries\color[rgb]{0.5,0.0,0.0},
	keywordstyle=[3]\bfseries\color[rgb]{0.127,0.427,0.514},
	keywordstyle=[4]\bfseries\color[rgb]{0.4,0.4,0.4},
	stringstyle=\color[rgb]{0.639,0.082,0.082},
}

\title{Innovations Report: Procedural Fur System}

\author{Joe Withers\thanks{e-mail:joewithers96@gmail.com}\\National Centre for Computer Animation}
\pdfauthor{Joe Withers}

\keywords{rendering}

\begin{document}

\maketitle

\begin{abstract}
This document serves to demonstrate the short paper format used for the Principles of Rendering unit. This submission should be no longer than 2 pages, inclusive of references and images, but excluding appendices and code listings. You should aim for between 5 and 8 references in your bibliography.
An abstract provides an executive summary of your short paper.
\end{abstract}

\section{Introduction} \label{sec:introduction}
Introduction Text.

\section{Method Overview} \label{sec:overview}
Method Text.

\section{Results} \label{sec:results}
Results Text.

\bibliographystyle{acmsiggraph}
\bibliography{references}

\end{document}

### 2018-03-01 docs/paper.tex
progress on docs

@@
 \begin{abstract}
-This document serves to demonstrate the short paper format used for the Principles of Rendering unit.
-This submission should be no longer than 2 pages, inclusive of references and images, but excluding appendices and code listings.
-You should aim for between 5 and 8 references in your bibliography.
-An abstract provides an executive summary of your short paper.
+For this project I developed a procedural fur system, with the aim of exploring the feasibility of offloading computation onto the GPU within artist tools.
+The final artefact is an application which serves to interface with the fur system API that I have developed.
+This report primarily documents the implementation of the API, as well as my findings.
 \end{abstract}
@@
 
-\section{Method Overview} \label{sec:overview}
+\subsection{Existing Solutions} \label{sec:existing}
+Existing Solutions.
+
+\section{Implementation} \label{sec:implementation}
 Method Text.
+
+\subsection{Resources} \label{sec:resources}
+I decided to develop the API side of the fur system using C++, primarly as it most commonly used when developing computationally heavy artist tools, but also because it is the language I am most comfortable using.
+I opted to use OpenGL over other APIs (CUDA, OpenCL) for offloading of computation onto the GPU, simply because my final artefact is a tool with a graphical interface, and OpenGL is capable of handling both arbitary computation using compute shaders and rendering of geometry.
+
+To handle the user interface I used the Qt framework within C++.
+Qt is commonly used for artist tools within visual effects as it is cross-platform, and applications can be configured to run within other applications that make use of it, such as Autodesk Maya.
+
+I wanted to include a node-graph style interface within my application, as it is commonly used within existing artist tools (Autodesk Maya, Unreal Engine), and would encourage modularity within my API.
+Qt does not natively provide this kind of interface, so I made use of NodeEditor \cite{Pinaev2017} an existing Qt-based library that provides this functionality.
 

### 2018-03-01 docs/paper.tex
progress on docs

@@
 \subsection{Existing Solutions} \label{sec:existing}
-Existing Solutions.
+Existing Solutions. (Xgen, houdini)
 
@@
 
+\subsection{Design} \label{sec:design}
+From looking at existing fur systems I was able determine that my simplified fur system would need to consist of the following components:
+\begin{itemize}
+	\item Distributors - These are responsible for the distribution of curves onto user specified geometry.
+User controllable parameters could include density (curve count), distribution pattern (random, uniform), and curve length.
+These parameters could potentially be controlled by texture inputs.
+	\item Operators - These are responsible for manipulation of the curves to achieve the desired look. Example operators could provide bending, clumping, or randomisation of input curves.
+	\item Renderers - These are responsible for the rendering of curves into the application viewport.
+Example renderers could provide mesh, cuves as lines, or curves as ribbons rendering functionality.
+User controllable parameters could provide controls for the shading model in use, as well as control of the base and tip widths when rendering curve ribbons for example.
+\end{itemize}
+
 \section{Results} \label{sec:results}

### 2018-03-02 include/ComputeShaderManager.h
added length control to distributor

@@ -53,7 +53,7 @@ class ComputeShaderManager
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief randomly distribute curves in a compute shader
 		//----------------------------------------------------------------------------------------------------------------------
-		void randomDistribution(unsigned int &meshSSBO, unsigned int &curvesSSBO, unsigned int curveCount, unsigned int faceCount, float meshArea);
+		void randomDistribution(unsigned int &meshSSBO, unsigned int &curvesSSBO, unsigned int curveCount, unsigned int faceCount, float meshArea, float length);
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief bend curves in a compute shader
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-02 include/widgets/RandomDistributorModelWidget.h

@@ -14,6 +14,7 @@
 #include <QtWidgets/QApplication>
 #include <QtWidgets/QButtonGroup>
 #include <QtWidgets/QCheckBox>
+#include <QtWidgets/QDoubleSpinBox>
 #include <QtWidgets/QFormLayout>
 #include <QtWidgets/QHeaderView>
 #include <QtWidgets/QLabel>
@@ -30,12 +31,14 @@ public:
     QSpinBox *countSpinBox;
     QLabel *modeLabel;
     QCheckBox *modeCheckBox;
+    QLabel *lengthLabel;
+    QDoubleSpinBox *lengthSpinBox;
 
     void setupUi(QWidget *RandomDistributorModelWidget)
     {
         if (RandomDistributorModelWidget->objectName().isEmpty())
             RandomDistributorModelWidget->setObjectName(QStringLiteral("RandomDistributorModelWidget"));
-        RandomDistributorModelWidget->resize(145, 64);
+        RandomDistributorModelWidget->resize(145, 88);
         formLayout = new QFormLayout(RandomDistributorModelWidget);
         formLayout->setObjectName(QStringLiteral("formLayout"));
         countLabel = new QLabel(RandomDistributorModelWidget);
@@ -54,12 +57,25 @@ public:
         modeLabel = new QLabel(RandomDistributorModelWidget);
         modeLabel->setObjectName(QStringLiteral("modeLabel"));
 
-        formLayout->setWidget(1, QFormLayout::LabelRole, modeLabel);
+        formLayout->setWidget(2, QFormLayout::LabelRole, modeLabel);
 
         modeCheckBox = new QCheckBox(RandomDistributorModelWidget);
         modeCheckBox->setObjectName(QStringLiteral("modeCheckBox"));
 
-        formLayout->setWidget(1, QFormLayout::FieldRole, modeCheckBox);
+        formLayout->setWidget(2, QFormLayout::FieldRole, modeCheckBox);
+
+        lengthLabel = new QLabel(RandomDistributorModelWidget);
+        lengthLabel->setObjectName(QStringLiteral("lengthLabel"));
+
+        formLayout->setWidget(1, QFormLayout::LabelRole, lengthLabel);
+
+        lengthSpinBox = new QDoubleSpinBox(RandomDistributorModelWidget);
+        lengthSpinBox->setObjectName(QStringLiteral("lengthSpinBox"));
+        lengthSpinBox->setMinimum(0.01);
[…]

### 2018-03-02 src/ComputeShaderManager.cpp

@@ -188,7 +188,7 @@ void ComputeShaderManager::copyCurvesSSBO(unsigned int src, unsigned int &dst)
 
 //----------------------------------------------------------------------------------------------------------------------
 
-void ComputeShaderManager::randomDistribution(unsigned int &meshSSBO, unsigned int &curvesSSBO, unsigned int curveCount, unsigned int faceCount, float meshArea)
+void ComputeShaderManager::randomDistribution(unsigned int &meshSSBO, unsigned int &curvesSSBO, unsigned int curveCount, unsigned int faceCount, float meshArea, float length)
 {
 	// if the SSBO exists, delete it (we need to resize it anyway)
 	if (curvesSSBO)
@@ -208,6 +208,7 @@ void ComputeShaderManager::randomDistribution(unsigned int &meshSSBO, unsigned i
 	glUniform1f(glGetUniformLocation(randomDistributionShader.getID(), "u_meshArea"), meshArea);
 	glUniform1ui(glGetUniformLocation(randomDistributionShader.getID(), "u_curveCount"), curveCount);
 	glUniform1ui(glGetUniformLocation(randomDistributionShader.getID(), "u_faceCount"), faceCount);
+	glUniform1f(glGetUniformLocation(randomDistributionShader.getID(), "u_length"), length);
 
 	glBindBufferBase(GL_SHADER_STORAGE_BUFFER, 0, meshSSBO);
 	glBindBufferBase(GL_SHADER_STORAGE_BUFFER, 1, curvesSSBO);

### 2018-03-02 src/RandomDistributorModel.cpp

@@ -15,6 +15,7 @@ RandomDistributorModel::RandomDistributorModel() :
 	// connect spinbox and checkbox
 	connect(m_ui->countSpinBox, SIGNAL(valueChanged(int)), this, SLOT(updateDistribution()));
 	connect(m_ui->modeCheckBox, SIGNAL(stateChanged(int)), this, SLOT(updateDistribution()));
+	connect(m_ui->lengthSpinBox, SIGNAL(valueChanged(double)), this, SLOT(updateDistribution()));
 
 	m_curves = Curves();
 	m_meshSSBOID = 0;
@@ -50,7 +51,8 @@ void RandomDistributorModel::distribute()
 			m_curvesSSBOID,
 			m_ui->countSpinBox->value(),
 			m_mesh.m_faces.size(),
-			m_totalMeshArea
+			m_totalMeshArea,
+			m_ui->lengthSpinBox->value()
 			);
 	}
 	else
@@ -105,7 +107,7 @@ void RandomDistributorModel::distribute()
 
 					for (int j = 0; j < 5; ++j)
 					{
-						curve.vertices[j] = randPos + (float(j) * randNorm * 0.01f);
+						curve.vertices[j] = randPos + ((float(j) / 5.0f) * randNorm * float(m_ui->lengthSpinBox->value()));
 					}
 
 					m_curves.m_curves.push_back(curve);

### 2018-03-02 include/AbstractCurveOperatorModel.h
added clump curves operator

@@ -57,13 +57,13 @@ class AbstractCurveOperatorModel : public NodeDataModel
 		//----------------------------------------------------------------------------------------------------------------------
 		virtual void operateCurves();
 		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief reset curves to input state
+		//----------------------------------------------------------------------------------------------------------------------
+		virtual void resetCurves();
+		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief event filter gets whether we clicked in the node
 		//----------------------------------------------------------------------------------------------------------------------
 		bool eventFilter(QObject *object, QEvent *event) override;
-		//----------------------------------------------------------------------------------------------------------------------
-		/// @brief reset curves to input state
-		//----------------------------------------------------------------------------------------------------------------------
-		void resetCurves();
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief curves object
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-02 include/ClumpCurveOperatorModel.h

@@ -1,73 +1,72 @@
-#ifndef COMPUTESHADERMANAGER_H
-#define COMPUTESHADERMANAGER_H
+#ifndef CLUMPCURVEOPERATORMODEL_H
+#define CLUMPCURVEOPERATORMODEL_H
 
-#include "Curves.h"
-#include "Mesh.h"
-#include "Shader.h"
+#include "AbstractCurveOperatorModel.h"
 
-#include <GL/glew.h>
+#include <QtWidgets/QDoubleSpinBox>
 
-class ComputeShaderManager
+class ClumpCurveOperatorModel : public AbstractCurveOperatorModel
 {
-	private:
-		//----------------------------------------------------------------------------------------------------------------------
-		/// @brief private dummy constructor for the singleton
-		//----------------------------------------------------------------------------------------------------------------------
-		ComputeShaderManager();
-		//----------------------------------------------------------------------------------------------------------------------
-		/// @brief our singleton instance
-		//----------------------------------------------------------------------------------------------------------------------
-		static ComputeShaderManager* m_instance;
+	Q_OBJECT
 
 	public:
 		//----------------------------------------------------------------------------------------------------------------------
-		/// @brief get the singleton instance
+		/// @brief constructor
 		//----------------------------------------------------------------------------------------------------------------------
-		static ComputeShaderManager* getInstance() {if (!m_instance){m_instance = new ComputeShaderManager();} return m_instance;}
+		ClumpCurveOperatorModel();
 		//----------------------------------------------------------------------------------------------------------------------
-		/// @brief dtor
+		/// @brief destructor
 		//----------------------------------------------------------------------------------------------------------------------
-		~ComputeShaderManager();
+		~ClumpCurveOperatorModel();
 		//----------------------------------------------------------------------------------------------------------------------
-		/// @brief delete all shaders
+		/// @brief overrides for NodeDataModel
 		//----------------------------------------------------------------------------------------------------------------------
-		void cleanUpAll();
+		QString caption() const override {return QString("Clump Curves");};
 		//----------------------------------------------------------------------------------------------------------------------
-		/// @brief recompile all shaders
+		QString name() const override {return QString("Clump Curves");};
 		//----------------------------------------------------------------------------------------------------------------------
-		void recompileShaders();
+		std::unique_ptr<NodeDataModel> clone() const override {return std::make_unique<ClumpCurveOperatorModel>();};
[…]

### 2018-03-02 include/ComputeShaderUtils.h

@@ -9,4 +9,13 @@ struct SSBO_Face
 	float area; float pad[3];
 };
 
+unsigned int getIndicesFromCurveSSBO(unsigned int _ssbo)
+{
+	GLint SSBOSize = 0;
+	glBindBuffer(GL_SHADER_STORAGE_BUFFER, _ssbo);
+	glGetBufferParameteriv(GL_SHADER_STORAGE_BUFFER, GL_BUFFER_SIZE, &SSBOSize);
+	glBindBuffer(GL_SHADER_STORAGE_BUFFER, 0);
+	return (SSBOSize / (sizeof(glm::vec4) * 5));
+}
+
 #endif // COMPUTESHADERUTILS_H

### 2018-03-02 src/AbstractCurveOperatorModel.cpp

@@ -2,17 +2,20 @@
 
 #include <QtCore/QEvent>
 
-#include "MeshData.h"
-
 #include "ComputeShaderManager.h"
 
 AbstractCurveOperatorModel::AbstractCurveOperatorModel()
 {
 	m_curves = Curves();
+	m_curvesSSBO = 0;
 	_nodeData = nullptr;
 }
 
-AbstractCurveOperatorModel::~AbstractCurveOperatorModel() {}
+AbstractCurveOperatorModel::~AbstractCurveOperatorModel()
+{
+	if (m_curvesSSBO)
+		glDeleteBuffers(1, &m_curvesSSBO);
+}
 
 unsigned int AbstractCurveOperatorModel::nPorts(PortType portType) const
 {

### 2018-03-02 src/BendCurveOperatorModel.cpp

@@ -5,16 +5,14 @@
 #include "CurvesData.h"
 
 BendCurveOperatorModel::BendCurveOperatorModel() :
+	AbstractCurveOperatorModel(),
 	m_spinbox(new QDoubleSpinBox())
 {
 	m_spinbox->setRange(0.0, 100.0);
 	m_spinbox->setDecimals(3);
 	m_spinbox->setSingleStep(0.01);
 	m_spinbox->setValue(0.1);
-
 	connect(m_spinbox, SIGNAL(valueChanged(double)), this, SLOT(updateCurves()));
-
-	m_curves = Curves();
 }
 
 BendCurveOperatorModel::~BendCurveOperatorModel()

### 2018-03-02 src/ClumpCurveOperatorModel.cpp

@@ -1,0 +1,124 @@
+#include "ClumpCurveOperatorModel.h"
+
+#include "ComputeShaderManager.h"
+
+#include "CurvesData.h"
+
+ClumpCurveOperatorModel::ClumpCurveOperatorModel() :
+	AbstractCurveOperatorModel()
+{
+	m_clumpCurves = Curves();
+	m_clumpCurvesSSBO = 0;
+	m_clumpNodeData = nullptr;
+}
+
+ClumpCurveOperatorModel::~ClumpCurveOperatorModel()
+{
+	if (m_clumpCurvesSSBO)
+		glDeleteBuffers(1, &m_clumpCurvesSSBO);
+}
+
+QString ClumpCurveOperatorModel::portCaption(PortType portType, PortIndex portIndex) const
+{
+	switch (portType)
+	{
+		case PortType::In:
+			if (portIndex == 0)
+				return QStringLiteral("Input Curves");
+			else if (portIndex == 1)
+				return QStringLiteral("Attractor Curves");
+			break;
+
+		case PortType::Out:
+				return QStringLiteral("Result");
+
+		default:
+			break;
+	}
+	return QString();
+}
+
+unsigned int ClumpCurveOperatorModel::nPorts(PortType portType) const
+{
+	unsigned int result = 1;
+
+	switch (portType)
+	{
+		case PortType::In:
+			result = 2;
+			break;
+
+		case PortType::Out:
+			result = 1;
+
+		default:
+			break;
+	}
+
+	return result;
+}
+
+NodeDataType ClumpCurveOperatorModel::dataType(PortType, PortIndex) const
+{
+	return CurvesData().type();
+}
+
+void ClumpCurveOperatorModel::setInData(std::shared_ptr<NodeData> nodeData, PortIndex portIndex)
+{
+	std::shared_ptr<CurvesData> data;
+
+	if(nodeData)
+	{
+		auto data = std::static_pointer_cast<CurvesData>(nodeData);
+		if (portIndex == 0)
+		{
+			_nodeData = data;
+		}
+		else if (portIndex == 1)
+		{
+			m_clumpNodeData = data;
+		}
+
+		if (_nodeData && m_clumpNodeData)
+			updateCurves();
+	}
+	else
+	{
+		m_curves = Curves();
+		m_clumpCurves = Curves();
+		// clear ssbo???
+		emit dataUpdated(0);
+	}
+}
+
[…]

### 2018-03-02 src/ComputeShaderManager.cpp

@@ -16,6 +16,7 @@ ComputeShaderManager::ComputeShaderManager()
 {
 	randomDistributionShader = Shader("shaders/compute/randomDistributionShader_comp.glsl");
 	bendCurvesShader = Shader("shaders/compute/bendCurvesShader_comp.glsl");
+	clumpCurvesShader = Shader("shaders/compute/clumpCurvesShader_comp.glsl");
 }
 
 //----------------------------------------------------------------------------------------------------------------------
@@ -31,6 +32,8 @@ void ComputeShaderManager::cleanUpAll()
 {
 	// delete compute shaders
 	randomDistributionShader.cleanUp();
+	bendCurvesShader.cleanUp();
+	clumpCurvesShader.cleanUp();
 }
 
 //----------------------------------------------------------------------------------------------------------------------
@@ -40,6 +43,7 @@ void ComputeShaderManager::recompileShaders()
 	cleanUpAll();
 	randomDistributionShader = Shader("shaders/compute/randomDistributionShader_comp.glsl");
 	bendCurvesShader = Shader("shaders/compute/bendCurvesShader_comp.glsl");
+	clumpCurvesShader = Shader("shaders/compute/clumpCurvesShader_comp.glsl");
 }
 
 //----------------------------------------------------------------------------------------------------------------------
@@ -69,23 +73,6 @@ void ComputeShaderManager::createCurvesSSBO(unsigned int &buffer, Curves _curves
 
 //----------------------------------------------------------------------------------------------------------------------
 
-// void ComputeShaderManager::createCurvesSSBO(unsigned int &buffer, unsigned int _count)
-// {
-// 	std::cout<<"creating an SSBO for curves\n";
-
-// 	if (buffer)
-// 		glDeleteBuffers(1, &buffer);
-
-// 	glGenBuffers(1, &buffer);
-// 	glBindBuffer(GL_SHADER_STORAGE_BUFFER, buffer);
-// 	// use glBufferData to allocate all 5*count vec3 or infer from curves object
-// 	// TODO: don't init with NULL
[…]

### 2018-03-02 src/MainWindow.cpp

@@ -6,6 +6,7 @@
 #include "BasicDistributorModel.h"
 #include "RandomDistributorModel.h"
 #include "BendCurveOperatorModel.h"
+#include "ClumpCurveOperatorModel.h"
 #include "ComputeShaderManager.h"
 #include "RenderableManager.h"
 
@@ -24,6 +25,7 @@ static std::shared_ptr<DataModelRegistry> registerDataModels()
 	ret->registerModel<BasicDistributorModel>("Distributors");
 	ret->registerModel<RandomDistributorModel>("Distributors");
 	ret->registerModel<BendCurveOperatorModel>("Curve Operators");
+	ret->registerModel<ClumpCurveOperatorModel>("Curve Operators");
 	return ret;
 }
 

### 2018-03-04 src/ClumpCurveOperatorModel.cpp
added clump envelope and rendering visibility toggles

@@ -5,17 +5,26 @@
 #include "CurvesData.h"
 
 ClumpCurveOperatorModel::ClumpCurveOperatorModel() :
-	AbstractCurveOperatorModel()
+	AbstractCurveOperatorModel(),
+	m_embedded(new QWidget()),
+	m_ui(new Ui::ClumpCurveOperatorModelWidget())
 {
+	// init clump curve variables
 	m_clumpCurves = Curves();
 	m_clumpCurvesSSBO = 0;
 	m_clumpNodeData = nullptr;
+	// setup ui
+	m_ui->setupUi(m_embedded);
+	// connect spinbox and checkbox
+	connect(m_ui->envelopeSpinBox, SIGNAL(valueChanged(double)), this, SLOT(updateCurves()));
 }
 
 ClumpCurveOperatorModel::~ClumpCurveOperatorModel()
 {
 	if (m_clumpCurvesSSBO)
 		glDeleteBuffers(1, &m_clumpCurvesSSBO);
+
+	delete m_ui;
 }
 
 QString ClumpCurveOperatorModel::portCaption(PortType portType, PortIndex portIndex) const
@@ -115,7 +124,7 @@ void ClumpCurveOperatorModel::operateCurves()
 	if (curveType == SSBO && clumpCurveType == SSBO)
 	{
 		std::cout<<"processing clump curves SSBO: "<<m_curvesSSBO<<" & "<<m_clumpCurvesSSBO<<"\n";
-		ComputeShaderManager::getInstance()->clumpCurvesOperator(m_curvesSSBO, m_clumpCurvesSSBO);
+		ComputeShaderManager::getInstance()->clumpCurvesOperator(m_curvesSSBO, m_clumpCurvesSSBO, m_ui->envelopeSpinBox->value());
 	}
 	else
 	{

### 2018-03-04 include/MainWindow.h
added splitter for dynamic layout

@@ -3,6 +3,7 @@
 
 #include <QMainWindow>
 #include <QKeyEvent>
+#include <QSplitter>
 
 #include "Viewport.h"
 
@@ -53,6 +54,10 @@ class MainWindow : public QMainWindow
 		/// @brief m_flowView viewport that displays the node graph
 		//----------------------------------------------------------------------------------------------------------------------
 		FlowView *m_flowView;
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief m_splitter splitter object for dynamic layout
+		//----------------------------------------------------------------------------------------------------------------------
+		QSplitter *m_splitter;
 };
 
 #endif // MAINWINDOW_H

### 2018-03-04 src/MainWindow.cpp

@@ -43,25 +43,30 @@ static void setStyle()
 
 MainWindow::MainWindow(QWidget *parent) :
 	QMainWindow(parent),
-	m_ui(new Ui::MainWindow)
+	m_ui(new Ui::MainWindow),
+	m_splitter(new QSplitter(parent))
 {
 	// setup ui
 	m_ui->setupUi(this);
 	// create viewport widget
 	m_gl = new Viewport(this);
-	m_ui->mainWindowGridLayout->addWidget(m_gl, 0, 0, 1, 1);
 	// create node graph widget
 	::setStyle();
 	m_nodes = new FlowScene(registerDataModels());
 	m_flowView = new FlowView(m_nodes);
-
-	m_ui->mainWindowGridLayout->addWidget(m_flowView, 0, 1, 1, 1);
+	// add widgets to splitter
+	m_splitter->addWidget(m_gl);
+	m_splitter->addWidget(m_flowView);
+	m_splitter->setSizes(QList<int>({INT_MAX, INT_MAX})); // set size of both splits to max so we fill 50/50
+	// add splitter to layout
+	m_ui->mainWindowGridLayout->addWidget(m_splitter, 0, 0, 1, 1);
 }
 
 MainWindow::~MainWindow()
 {
 	delete m_ui;
 	delete m_gl;
+	delete m_splitter;
 }
 
 void MainWindow::keyPressEvent(QKeyEvent *_event)

### 2018-03-04 src/ComputeShaderManager.cpp
added seed to randomness that relies on the curve SSBO ID being unique

@@ -193,6 +193,7 @@ void ComputeShaderManager::randomDistribution(unsigned int &meshSSBO, unsigned i
 	glUniform1ui(glGetUniformLocation(randomDistributionShader.getID(), "u_curveCount"), curveCount);
 	glUniform1ui(glGetUniformLocation(randomDistributionShader.getID(), "u_faceCount"), faceCount);
 	glUniform1f(glGetUniformLocation(randomDistributionShader.getID(), "u_length"), length);
+	glUniform1ui(glGetUniformLocation(randomDistributionShader.getID(), "u_seed"), curvesSSBO);
 
 	glBindBufferBase(GL_SHADER_STORAGE_BUFFER, 0, meshSSBO);
 	glBindBufferBase(GL_SHADER_STORAGE_BUFFER, 1, curvesSSBO);

### 2018-03-05 docs/paper.tex
more work on docs

@@
 I wanted to include a node-graph style interface within my application, as it is commonly used within existing artist tools (Autodesk Maya, Unreal Engine), and would encourage modularity within my API.
-Qt does not natively provide this kind of interface, so I made use of NodeEditor \cite{Pinaev2017} an existing Qt-based library that provides this functionality.
+Qt does not natively provide this kind of interface, so I made use of NodeEditor \cite{Pinaev2017}, an existing Qt-based library that provides this functionality.
 
@@
 From looking at existing fur systems I was able determine that my simplified fur system would need to consist of the following components:
+
 \begin{itemize}
+	\item Geometry Loaders - These are responsible for loading user specified geometry. Example Geometry Loaders could allow for the parsing of Wavefront OBJ files.
 	\item Distributors - These are responsible for the distribution of curves onto user specified geometry.
@@
 
+Whilst the list of components specified above provide the required functionality for a standalone tool, they do not provide any functionality for exporting fur for use in a broader pipeline.
+I chose to omit this functionality to limit the scope of this assignment, but the ability to export hair curves as alembic would be desirable, as alembic is a widely adopted format that allows for the storage of multiple different types of curve geometry.
+Another candidate file type would be RenderMan's RIB format, which would allow for direct rendering of the hair curves in any RenderMan compliant renderer.
+
 \section{Results} \label{sec:results}

### 2018-03-06 docs/paper.tex
more docs

@@
 I decided to develop the API side of the fur system using C++, primarly as it most commonly used when developing computationally heavy artist tools, but also because it is the language I am most comfortable using.
-I opted to use OpenGL over other APIs (CUDA, OpenCL) for offloading of computation onto the GPU, simply because my final artefact is a tool with a graphical interface, and OpenGL is capable of handling both arbitary computation using compute shaders and rendering of geometry.
+I opted to use OpenGL over other APIs (CUDA, OpenCL) for offloading of computation onto the GPU, simply because my final artefact is a tool with a graphical interface, and OpenGL is capable of handling both arbitary computation using compute shaders and rendering of geometry simultaneously.
 
@@
 
+\subsection{Development} \label{sec:development}
+
+write about obj loading
+
+write about distributing https://stackoverflow.com/questions/9294316/distribute-points-on-mesh-according-to-density
+
+write about GPU
+
 \section{Results} \label{sec:results}
 Results Text.
+
+\subsection{Performance} \label{sec:performance}
+
+insert stats comparing CPU/GPU compute
 

### 2018-03-07 docs/paper.tex
more on docs & test from alin
more on docs

@@
 
+\subsection{User Testing} \label{sec:usertesting}
+Near the end of the development I was able to get friend and fellow student Alin Bolcas to test the system and provide feedback from an artist's point of view; Alin is an experienced modelling and look development artist who uses a variety of different software packages to create his work, so getting his feedback on the usability of my tool proved extremely useful.
+
+His first comments were in regards to how the node graph handles rendering of meshes and curves; he found it quite unintuitive to manually create rendering nodes for each new curves object that needed to be drawn.
+He suggested that a node for the combination of curves should be added, which would consolidate multiple curve objects into a single output for simplified rendering.
+
+He noted that the lack of a node that applies noise to curves proved problematic during his testing; this kind of node is fundamental to achieving appealing fur visuals.
+He liked the behaviour of the clumping node, as this functionality is also very common and highly used within fur systems, though visual quality was somewhat compromised as my current implementation does not preserve the initial curve length.
+Adding a control in the distributor that would allow for the variation of curve length was also suggested as something that would improve the quality of the visual output.
+
 \section{Results} \label{sec:results}
@@
 
+\section{Conclusion} \label{sec:conclusion}
+
 \bibliographystyle{acmsiggraph}

### 2018-03-07 docs/paper.tex
more on docs

@@
 
+His final comments were in regards to the user interface, suggesting that I make controls similar to their counterparts in existing software wherever possible.
+Changing the camera controls to match those in Maya was suggested as it would provide a familiar user experience when manipulating the viewport.
+The ability to create new nodes by pressing the tab key as opposed to right click was also suggested, as this is the standard in node editors, such as those in Maya and Nuke.
+
 \section{Results} \label{sec:results}

### 2018-03-08 include/Curves.h
added length member to curves

@@ -18,6 +18,10 @@ struct Curve
 	/// @brief the curve vertices
 	//----------------------------------------------------------------------------------------------------------------------
 	glm::vec3 vertices[5]; // hard coded 5 vertices
+	//----------------------------------------------------------------------------------------------------------------------
+	/// @brief the resting length of this curve
+	//----------------------------------------------------------------------------------------------------------------------
+	float length;
 };
 
 struct Curves

### 2018-03-08 src/ComputeShaderManager.cpp

@@ -65,8 +65,10 @@ void ComputeShaderManager::createCurvesSSBO(unsigned int &buffer, Curves _curves
 		{
 			curvesData.push_back(glm::vec4(curve.vertices[i], 0.0));
 		}
+		// push back the length
+		curvesData.push_back(glm::vec4(curve.length, 0.0, 0.0, 0.0));
 	}
-	// use glBufferData to allocate all 5*count vec3 or infer from curves object
+	// use glBufferData to allocate all 6 * count vec3 or infer from curves object
 	glBufferData(GL_SHADER_STORAGE_BUFFER, curvesData.size() * sizeof(glm::vec4(0.0)), &curvesData[0], GL_DYNAMIC_COPY);
 	glBindBuffer(GL_SHADER_STORAGE_BUFFER, 0);
 }
@@ -182,7 +184,7 @@ void ComputeShaderManager::randomDistribution(unsigned int &meshSSBO, unsigned i
 	glGenBuffers(1, &curvesSSBO);
 	glBindBuffer(GL_SHADER_STORAGE_BUFFER, curvesSSBO);
 	std::vector<glm::vec4> curvesData;
-	for (unsigned int i = 0; i < (curveCount * 5); ++i)
+	for (unsigned int i = 0; i < (curveCount * 6); ++i) // 6 = 5 verts + 1 for length
 		curvesData.push_back(glm::vec4(0.0));
 	glBufferData(GL_SHADER_STORAGE_BUFFER, curveCount * 5 * sizeof(glm::vec4), &curvesData[0], GL_DYNAMIC_COPY);
 	glBindBuffer(GL_SHADER_STORAGE_BUFFER, 0);

### 2018-03-08 src/CurvesRenderable.cpp

@@ -38,7 +38,7 @@ void CurvesRenderable::generate()
 			glGetBufferParameteriv(GL_SHADER_STORAGE_BUFFER, GL_BUFFER_SIZE, &SSBOSize);
 			glBindBuffer(GL_SHADER_STORAGE_BUFFER, 0);
 			std::cout<<"SSBO size : "<<SSBOSize<<"\n";
-			m_indices = SSBOSize / (sizeof(glm::vec4) * 5);
+			m_indices = SSBOSize / (sizeof(glm::vec4) * 6);
 			std::cout<<"m_indices from SSBO size : "<<m_indices<<"\n";
 			break;
 	}

### 2018-03-08 src/RandomDistributorModel.cpp

@@ -109,9 +109,9 @@ void RandomDistributorModel::distribute()
 					{
 						curve.vertices[j] = randPos + ((float(j) / 5.0f) * randNorm * float(m_ui->lengthSpinBox->value()));
 					}
+					curve.length = m_ui->lengthSpinBox->value();
 
 					m_curves.m_curves.push_back(curve);
-
 					break;
 				}
 			}

### 2018-03-08 include/ComputeShaderManager.h
added varitation control to distributor

@@ -53,7 +53,7 @@ class ComputeShaderManager
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief randomly distribute curves in a compute shader
 		//----------------------------------------------------------------------------------------------------------------------
-		void randomDistribution(unsigned int &meshSSBO, unsigned int &curvesSSBO, unsigned int curveCount, unsigned int faceCount, float meshArea, float length);
+		void randomDistribution(unsigned int &meshSSBO, unsigned int &curvesSSBO, unsigned int curveCount, unsigned int faceCount, float meshArea, float length, float variation);
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief bend curves in a compute shader
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-08 include/ComputeShaderUtils.h

@@ -15,7 +15,7 @@ unsigned int getIndicesFromCurveSSBO(unsigned int _ssbo)
 	glBindBuffer(GL_SHADER_STORAGE_BUFFER, _ssbo);
 	glGetBufferParameteriv(GL_SHADER_STORAGE_BUFFER, GL_BUFFER_SIZE, &SSBOSize);
 	glBindBuffer(GL_SHADER_STORAGE_BUFFER, 0);
-	return (SSBOSize / (sizeof(glm::vec4) * 5));
+	return (SSBOSize / (sizeof(glm::vec4) * 6));
 }
 
 #endif // COMPUTESHADERUTILS_H

### 2018-03-08 include/widgets/RandomDistributorModelWidget.h

@@ -33,12 +33,14 @@ public:
     QCheckBox *modeCheckBox;
     QLabel *lengthLabel;
     QDoubleSpinBox *lengthSpinBox;
+    QLabel *variationLabel;
+    QDoubleSpinBox *variationSpinBox;
 
     void setupUi(QWidget *RandomDistributorModelWidget)
     {
         if (RandomDistributorModelWidget->objectName().isEmpty())
             RandomDistributorModelWidget->setObjectName(QStringLiteral("RandomDistributorModelWidget"));
-        RandomDistributorModelWidget->resize(145, 88);
+        RandomDistributorModelWidget->resize(152, 114);
         formLayout = new QFormLayout(RandomDistributorModelWidget);
         formLayout->setObjectName(QStringLiteral("formLayout"));
         countLabel = new QLabel(RandomDistributorModelWidget);
@@ -57,13 +59,13 @@ public:
         modeLabel = new QLabel(RandomDistributorModelWidget);
         modeLabel->setObjectName(QStringLiteral("modeLabel"));
 
-        formLayout->setWidget(2, QFormLayout::LabelRole, modeLabel);
+        formLayout->setWidget(3, QFormLayout::LabelRole, modeLabel);
 
         modeCheckBox = new QCheckBox(RandomDistributorModelWidget);
         modeCheckBox->setObjectName(QStringLiteral("modeCheckBox"));
         modeCheckBox->setChecked(true);
 
-        formLayout->setWidget(2, QFormLayout::FieldRole, modeCheckBox);
+        formLayout->setWidget(3, QFormLayout::FieldRole, modeCheckBox);
 
         lengthLabel = new QLabel(RandomDistributorModelWidget);
         lengthLabel->setObjectName(QStringLiteral("lengthLabel"));
@@ -78,6 +80,20 @@ public:
 
         formLayout->setWidget(1, QFormLayout::FieldRole, lengthSpinBox);
 
+        variationLabel = new QLabel(RandomDistributorModelWidget);
+        variationLabel->setObjectName(QStringLiteral("variationLabel"));
+
+        formLayout->setWidget(2, QFormLayout::LabelRole, variationLabel);
+
+        variationSpinBox = new QDoubleSpinBox(RandomDistributorModelWidget);
+        variationSpinBox->setObjectName(QStringLiteral("variationSpinBox"));
+        variationSpinBox->setMinimum(0);
+        variationSpinBox->setMaximum(1);
[…]

### 2018-03-08 src/ComputeShaderManager.cpp

@@ -174,7 +174,7 @@ void ComputeShaderManager::copyCurvesSSBO(unsigned int src, unsigned int &dst)
 
 //----------------------------------------------------------------------------------------------------------------------
 
-void ComputeShaderManager::randomDistribution(unsigned int &meshSSBO, unsigned int &curvesSSBO, unsigned int curveCount, unsigned int faceCount, float meshArea, float length)
+void ComputeShaderManager::randomDistribution(unsigned int &meshSSBO, unsigned int &curvesSSBO, unsigned int curveCount, unsigned int faceCount, float meshArea, float length, float variation)
 {
 	// if the SSBO exists, delete it (we need to resize it anyway)
 	if (curvesSSBO)
@@ -186,7 +186,7 @@ void ComputeShaderManager::randomDistribution(unsigned int &meshSSBO, unsigned i
 	std::vector<glm::vec4> curvesData;
 	for (unsigned int i = 0; i < (curveCount * 6); ++i) // 6 = 5 verts + 1 for length
 		curvesData.push_back(glm::vec4(0.0));
-	glBufferData(GL_SHADER_STORAGE_BUFFER, curveCount * 5 * sizeof(glm::vec4), &curvesData[0], GL_DYNAMIC_COPY);
+	glBufferData(GL_SHADER_STORAGE_BUFFER, curveCount * 6 * sizeof(glm::vec4), &curvesData[0], GL_DYNAMIC_COPY);
 	glBindBuffer(GL_SHADER_STORAGE_BUFFER, 0);
 
 	// use shader & send uniforms
@@ -195,6 +195,7 @@ void ComputeShaderManager::randomDistribution(unsigned int &meshSSBO, unsigned i
 	glUniform1ui(glGetUniformLocation(randomDistributionShader.getID(), "u_curveCount"), curveCount);
 	glUniform1ui(glGetUniformLocation(randomDistributionShader.getID(), "u_faceCount"), faceCount);
 	glUniform1f(glGetUniformLocation(randomDistributionShader.getID(), "u_length"), length);
+	glUniform1f(glGetUniformLocation(randomDistributionShader.getID(), "u_variation"), variation);
 	glUniform1ui(glGetUniformLocation(randomDistributionShader.getID(), "u_seed"), curvesSSBO);
 
 	glBindBufferBase(GL_SHADER_STORAGE_BUFFER, 0, meshSSBO);

### 2018-03-08 src/RandomDistributorModel.cpp

@@ -16,6 +16,7 @@ RandomDistributorModel::RandomDistributorModel() :
 	connect(m_ui->countSpinBox, SIGNAL(valueChanged(int)), this, SLOT(updateDistribution()));
 	connect(m_ui->modeCheckBox, SIGNAL(stateChanged(int)), this, SLOT(updateDistribution()));
 	connect(m_ui->lengthSpinBox, SIGNAL(valueChanged(double)), this, SLOT(updateDistribution()));
+	connect(m_ui->variationSpinBox, SIGNAL(valueChanged(double)), this, SLOT(updateDistribution()));
 
 	m_curves = Curves();
 	m_meshSSBOID = 0;
@@ -45,14 +46,15 @@ void RandomDistributorModel::distribute()
 {
 	if (m_ui->modeCheckBox->isChecked()) // GPU mode checked
 	{
-		// this function needs to create / resize m_curvesSSBOID
+		// this function needs to create & resize m_curvesSSBOID
 		ComputeShaderManager::getInstance()->randomDistribution(
 			m_meshSSBOID,
 			m_curvesSSBOID,
 			m_ui->countSpinBox->value(),
 			m_mesh.m_faces.size(),
 			m_totalMeshArea,
-			m_ui->lengthSpinBox->value()
+			m_ui->lengthSpinBox->value(),
+			m_ui->variationSpinBox->value()
 			);
 	}
 	else

### 2018-03-08 include/ComputeShaderManager.h
added length preservation to clump

@@ -61,7 +61,7 @@ class ComputeShaderManager
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief clump curves in a compute shader
 		//----------------------------------------------------------------------------------------------------------------------
-		void clumpCurvesOperator(unsigned int curvesSSBO, unsigned int clumpCurvesSSBO, float envelope);
+		void clumpCurvesOperator(unsigned int curvesSSBO, unsigned int clumpCurvesSSBO, float envelope, bool preserveLength);
 
 	private:
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-08 include/widgets/ClumpCurveOperatorModelWidget.h

@@ -13,6 +13,7 @@
 #include <QtWidgets/QAction>
 #include <QtWidgets/QApplication>
 #include <QtWidgets/QButtonGroup>
+#include <QtWidgets/QCheckBox>
 #include <QtWidgets/QDoubleSpinBox>
 #include <QtWidgets/QFormLayout>
 #include <QtWidgets/QHeaderView>
@@ -27,12 +28,14 @@ public:
     QFormLayout *formLayout;
     QLabel *envelopeLabel;
     QDoubleSpinBox *envelopeSpinBox;
+    QLabel *preserveLengthLabel;
+    QCheckBox *preserveLengthCheckBox;
 
     void setupUi(QWidget *ClumpCurveOperatorModelWidget)
     {
         if (ClumpCurveOperatorModelWidget->objectName().isEmpty())
             ClumpCurveOperatorModelWidget->setObjectName(QStringLiteral("ClumpCurveOperatorModelWidget"));
-        ClumpCurveOperatorModelWidget->resize(145, 39);
+        ClumpCurveOperatorModelWidget->resize(176, 56);
         formLayout = new QFormLayout(ClumpCurveOperatorModelWidget);
         formLayout->setObjectName(QStringLiteral("formLayout"));
         envelopeLabel = new QLabel(ClumpCurveOperatorModelWidget);
@@ -49,6 +52,17 @@ public:
 
         formLayout->setWidget(0, QFormLayout::FieldRole, envelopeSpinBox);
 
+        preserveLengthLabel = new QLabel(ClumpCurveOperatorModelWidget);
+        preserveLengthLabel->setObjectName(QStringLiteral("preserveLengthLabel"));
+
+        formLayout->setWidget(1, QFormLayout::LabelRole, preserveLengthLabel);
+
+        preserveLengthCheckBox = new QCheckBox(ClumpCurveOperatorModelWidget);
+        preserveLengthCheckBox->setObjectName(QStringLiteral("preserveLengthCheckBox"));
+        preserveLengthCheckBox->setChecked(true);
+
+        formLayout->setWidget(1, QFormLayout::FieldRole, preserveLengthCheckBox);
+
 
         retranslateUi(ClumpCurveOperatorModelWidget);
 
@@ -59,6 +73,8 @@ public:
     {
         ClumpCurveOperatorModelWidget->setWindowTitle(QApplication::translate("ClumpCurveOperatorModelWidget", "Form", Q_NULLPTR));
[…]

### 2018-03-08 src/ClumpCurveOperatorModel.cpp

@@ -17,6 +17,7 @@ ClumpCurveOperatorModel::ClumpCurveOperatorModel() :
 	m_ui->setupUi(m_embedded);
 	// connect spinbox and checkbox
 	connect(m_ui->envelopeSpinBox, SIGNAL(valueChanged(double)), this, SLOT(updateCurves()));
+	connect(m_ui->preserveLengthCheckBox, SIGNAL(stateChanged(int)), this, SLOT(updateCurves()));
 }
 
 ClumpCurveOperatorModel::~ClumpCurveOperatorModel()
@@ -124,7 +125,7 @@ void ClumpCurveOperatorModel::operateCurves()
 	if (curveType == SSBO && clumpCurveType == SSBO)
 	{
 		std::cout<<"processing clump curves SSBO: "<<m_curvesSSBO<<" & "<<m_clumpCurvesSSBO<<"\n";
-		ComputeShaderManager::getInstance()->clumpCurvesOperator(m_curvesSSBO, m_clumpCurvesSSBO, m_ui->envelopeSpinBox->value());
+		ComputeShaderManager::getInstance()->clumpCurvesOperator(m_curvesSSBO, m_clumpCurvesSSBO, m_ui->envelopeSpinBox->value(), m_ui->preserveLengthCheckBox->isChecked());
 	}
 	else
 	{

### 2018-03-08 src/ComputeShaderManager.cpp

@@ -226,7 +226,7 @@ void ComputeShaderManager::bendCurvesOperator(unsigned int curvesSSBO, glm::vec3
 
 //----------------------------------------------------------------------------------------------------------------------
 
-void ComputeShaderManager::clumpCurvesOperator(unsigned int curvesSSBO, unsigned int clumpCurvesSSBO, float envelope)
+void ComputeShaderManager::clumpCurvesOperator(unsigned int curvesSSBO, unsigned int clumpCurvesSSBO, float envelope, bool preserveLength)
 {
 	if (envelope == 0.0) // if envelope 0, dont bother!
 		return;
@@ -246,10 +246,11 @@ void ComputeShaderManager::clumpCurvesOperator(unsigned int curvesSSBO, unsigned
 	glUniform1ui(glGetUniformLocation(clumpCurvesShader.getID(), "u_curveCount"), curveCount);
 	glUniform1ui(glGetUniformLocation(clumpCurvesShader.getID(), "u_clumpCurveCount"), clumpCurveCount);
 	glUniform1f(glGetUniformLocation(clumpCurvesShader.getID(), "u_envelope"), envelope);
+	glUniform1i(glGetUniformLocation(clumpCurvesShader.getID(), "u_preserveLength"), preserveLength);
 
-	glUniform1ui(glGetUniformLocation(clumpCurvesShader.getID(), "u_mode"), 0);
-	glDispatchCompute((int(clumpCurveCount) / 128) + 1, 1, 1);
-	glMemoryBarrier(GL_ALL_BARRIER_BITS);
+	// glUniform1ui(glGetUniformLocation(clumpCurvesShader.getID(), "u_mode"), 0);
+	// glDispatchCompute((int(clumpCurveCount) / 128) + 1, 1, 1);
+	// glMemoryBarrier(GL_ALL_BARRIER_BITS);
 	glUniform1ui(glGetUniformLocation(clumpCurvesShader.getID(), "u_mode"), 1);
 	glDispatchCompute((int(curveCount) / 128) + 1, 1, 1);
 }

### 2018-03-08 src/OrbitingCamera.cpp

@@ -65,7 +65,7 @@ void OrbitingCamera::processMouseMovement(float _xoffset, float _yoffset, bool _
 void OrbitingCamera::processMouseScroll(float _yoffset) noexcept
 {
 	if ((m_radius + _yoffset) >= 0.0) // ensure radius isn't negative
-		m_radius += _yoffset;
+		m_radius += _yoffset * 0.1f;
 	calculateOrbitPos();
 	updateCameraVectors();
 }

### 2018-03-09 include/ComputeShaderManager.h
added noise operator

@@ -59,6 +59,10 @@ class ComputeShaderManager
 		//----------------------------------------------------------------------------------------------------------------------
 		void bendCurvesOperator(unsigned int curvesSSBO, glm::vec3 direction, float intensity);
 		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief noise curves in a compute shader
+		//----------------------------------------------------------------------------------------------------------------------
+		void noiseCurvesOperator(unsigned int curvesSSBO, float intensity);
+		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief clump curves in a compute shader
 		//----------------------------------------------------------------------------------------------------------------------
 		void clumpCurvesOperator(unsigned int curvesSSBO, unsigned int clumpCurvesSSBO, float envelope, bool preserveLength);
@@ -67,7 +71,7 @@ class ComputeShaderManager
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief our compute shaders
 		//----------------------------------------------------------------------------------------------------------------------
-		Shader randomDistributionShader, bendCurvesShader, clumpCurvesShader;
+		Shader randomDistributionShader, bendCurvesShader, noiseCurvesShader, clumpCurvesShader;
 };
 
 #endif // COMPUTESHADERMANAGER_H

### 2018-03-09 include/NoiseCurveOperatorModel.h

@@ -1,0 +1,47 @@
+#ifndef NOISECURVEOPERATORMODEL_H
+#define NOISECURVEOPERATORMODEL_H
+
+#include "AbstractCurveOperatorModel.h"
+
+#include <QtWidgets/QDoubleSpinBox>
+
+class NoiseCurveOperatorModel : public AbstractCurveOperatorModel
+{
+	Q_OBJECT
+
+	public:
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief constructor
+		//----------------------------------------------------------------------------------------------------------------------
+		NoiseCurveOperatorModel();
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief destructor
+		//----------------------------------------------------------------------------------------------------------------------
+		~NoiseCurveOperatorModel();
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief overrides for NodeDataModel
+		//----------------------------------------------------------------------------------------------------------------------
+		QString caption() const override {return QString("Noise Curves");};
+		//----------------------------------------------------------------------------------------------------------------------
+		QString name() const override {return QString("Noise Curves");};
+		//----------------------------------------------------------------------------------------------------------------------
+		std::unique_ptr<NodeDataModel> clone() const override {return std::make_unique<NoiseCurveOperatorModel>();};
+		//----------------------------------------------------------------------------------------------------------------------
+		virtual QString modelName() const {return QString("Noise Operated Curves");};
+		//----------------------------------------------------------------------------------------------------------------------
+		QWidget* embeddedWidget() override {return m_spinbox;};
+		//----------------------------------------------------------------------------------------------------------------------
+
+	private:
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief manipulate curves
+		//----------------------------------------------------------------------------------------------------------------------
+		void operateCurves() override;
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief spinbox to control intensity (temporary)
+		//----------------------------------------------------------------------------------------------------------------------
+		QDoubleSpinBox* m_spinbox;
+
+};
+
+#endif // NOISECURVEOPERATORMODEL_H

### 2018-03-09 src/ComputeShaderManager.cpp

@@ -16,6 +16,7 @@ ComputeShaderManager::ComputeShaderManager()
 {
 	randomDistributionShader = Shader("shaders/compute/randomDistributionShader_comp.glsl");
 	bendCurvesShader = Shader("shaders/compute/bendCurvesShader_comp.glsl");
+	noiseCurvesShader = Shader("shaders/compute/noiseCurvesShader_comp.glsl");
 	clumpCurvesShader = Shader("shaders/compute/clumpCurvesShader_comp.glsl");
 }
 
@@ -33,6 +34,7 @@ void ComputeShaderManager::cleanUpAll()
 	// delete compute shaders
 	randomDistributionShader.cleanUp();
 	bendCurvesShader.cleanUp();
+	noiseCurvesShader.cleanUp();
 	clumpCurvesShader.cleanUp();
 }
 
@@ -43,6 +45,7 @@ void ComputeShaderManager::recompileShaders()
 	cleanUpAll();
 	randomDistributionShader = Shader("shaders/compute/randomDistributionShader_comp.glsl");
 	bendCurvesShader = Shader("shaders/compute/bendCurvesShader_comp.glsl");
+	noiseCurvesShader = Shader("shaders/compute/noiseCurvesShader_comp.glsl");
 	clumpCurvesShader = Shader("shaders/compute/clumpCurvesShader_comp.glsl");
 }
 
@@ -218,6 +221,22 @@ void ComputeShaderManager::bendCurvesOperator(unsigned int curvesSSBO, glm::vec3
 	glUniform1f(glGetUniformLocation(bendCurvesShader.getID(), "u_intensity"), intensity);
 	glUniform1ui(glGetUniformLocation(bendCurvesShader.getID(), "u_curveCount"), curveCount);
 	glUniform3fv(glGetUniformLocation(bendCurvesShader.getID(), "u_direction"), 1, glm::value_ptr(direction));
+
+	glBindBufferBase(GL_SHADER_STORAGE_BUFFER, 0, curvesSSBO);
+
+	glDispatchCompute((int(curveCount) / 128) + 1, 1, 1);
+}
+
+//----------------------------------------------------------------------------------------------------------------------
+
+void ComputeShaderManager::noiseCurvesOperator(unsigned int curvesSSBO, float intensity)
+{
+	unsigned int curveCount = getIndicesFromCurveSSBO(curvesSSBO);
+
+	// use shader & send uniforms
[…]

### 2018-03-09 src/MainWindow.cpp

@@ -6,6 +6,7 @@
 #include "BasicDistributorModel.h"
 #include "RandomDistributorModel.h"
 #include "BendCurveOperatorModel.h"
+#include "NoiseCurveOperatorModel.h"
 #include "ClumpCurveOperatorModel.h"
 #include "ComputeShaderManager.h"
 #include "RenderableManager.h"
@@ -25,6 +26,7 @@ static std::shared_ptr<DataModelRegistry> registerDataModels()
 	ret->registerModel<BasicDistributorModel>("Distributors");
 	ret->registerModel<RandomDistributorModel>("Distributors");
 	ret->registerModel<BendCurveOperatorModel>("Curve Operators");
+	ret->registerModel<NoiseCurveOperatorModel>("Curve Operators");
 	ret->registerModel<ClumpCurveOperatorModel>("Curve Operators");
 	return ret;
 }

### 2018-03-09 src/NoiseCurveOperatorModel.cpp

@@ -1,0 +1,42 @@
+#include "NoiseCurveOperatorModel.h"
+
+#include "ComputeShaderManager.h"
+
+#include "CurvesData.h"
+
+NoiseCurveOperatorModel::NoiseCurveOperatorModel() :
+	AbstractCurveOperatorModel(),
+	m_spinbox(new QDoubleSpinBox())
+{
+	m_spinbox->setRange(0.0, 1.0);
+	m_spinbox->setDecimals(2);
+	m_spinbox->setSingleStep(0.1);
+	m_spinbox->setValue(1.0);
+	connect(m_spinbox, SIGNAL(valueChanged(double)), this, SLOT(updateCurves()));
+}
+
+NoiseCurveOperatorModel::~NoiseCurveOperatorModel()
+{
+	// delete m_spinbox;
+}
+
+void NoiseCurveOperatorModel::operateCurves()
+{
+	float intensity = m_spinbox->value();
+
+	if (intensity == 0.0)
+		return;
+
+	switch (std::static_pointer_cast<CurvesData>(_nodeData)->curveType())
+	{
+		case CPU:
+			std::cout<<"processing noise curves not possible on CPU\n";
+			break;
+
+		case SSBO:
+			std::cout<<"processing noise curves SSBO\n";
+			ComputeShaderManager::getInstance()->noiseCurvesOperator(m_curvesSSBO, intensity);
+			break;
+	}
+
+}

### 2018-03-12 docs/paper.tex
more on docs

@@
 \subsection{Existing Solutions} \label{sec:existing}
-Existing Solutions. (Xgen, houdini)
+Whilst many larger companies opt to develop their own software for handling fur, such as MPC's Furtility \cite{furtility}, there are a number of commercially available solutions that are also widely used in production.
+The most popular of these is XGen, originally developed by Walt Disney Animation Studios, which has been included in Autodesk's Maya since 2014.
+XGen uses a 'Collection' and 'Description' paradigm; Descriptions store user controlled parameters which affect fur primitives (curves for example) on a select area of a mesh, and Collections act as a container for these Descriptions.
+There are two primary methods offered for distributing primitives on a mesh, Guide Curves can be used to manually control the placement and shape of primitives by placing 'Guides' on the mesh surface, whereas 'Groomable Splines' randomly distirbutes splines over the mesh surface which can be manipulated in real time using viewport manipulators that mimic the brushing and styling of fur.
+In both modes, 'modifiers' can be added to the modifier stack to further adjust the visual outcome, adding features such as clumping, noise, and coiling.
+Attributes within these modifiers can be further controlled using expressions or texture maps.
+
+Another commercially available solution is Peregrine*Labs' Yeti \cite{yeti}, which is also offered as a plugin within Autodesk's Maya.
+Yeti offers a more procedural approach, giving the user a node graph interface in which to construct the desired look from individual components.
+For example, the most simple usage requires the user to first create a 'scatter' node to distribute points on the mesh surface, and then combine it with a 'grow' node to extrude curves from these points.
+Whilst at first this may seem unintuitive when compared to XGen, it offers greater flexibility and avoids the pitfalls associated with a destructive workflow.
+
+Offerering even even more flexibility, recent versions of SideFX's Houdini include hair and fur operator nodes, which create predefined node networks for the generation of hair and fur.
+These networks use standard Houdini nodes internally, so experienced Houdini users are intuitively able to create the visual outcome they desire, using tools they already know.
+However, it may be quite unintuitive for asset artists (those who would be using the tool, Groom artists?), who are not used to Houdini's heavily procedural workflow.
 
@@
 \subsection{User Testing} \label{sec:usertesting}
-Near the end of the development I was able to get friend and fellow student Alin Bolcas to test the system and provide feedback from an artist's point of view; Alin is an experienced modelling and look development artist who uses a variety of different software packages to create his work, so getting his feedback on the usability of my tool proved extremely useful.
+Near the end of the development I was able to convince friend and fellow student Alin Bolcas to test the system and provide feedback from an artist's point of view; Alin is an experienced modelling and look development artist who uses a variety of different software packages to create his work, so getting his feedback on the usability of my tool proved extremely useful.
 

### 2018-03-12 src/MainWindow.cpp
added menubar actions

@@ -62,6 +62,12 @@ MainWindow::MainWindow(QWidget *parent) :
 	m_splitter->setSizes(QList<int>({INT_MAX, INT_MAX})); // set size of both splits to max so we fill 50/50
 	// add splitter to layout
 	m_ui->mainWindowGridLayout->addWidget(m_splitter, 0, 0, 1, 1);
+	// create menubar actions
+	auto saveAction = m_ui->menubar->addAction("Save..");
+	auto loadAction = m_ui->menubar->addAction("Load..");
+	// connect signals
+	QObject::connect(saveAction, &QAction::triggered, m_nodes, &FlowScene::save);
+	QObject::connect(loadAction, &QAction::triggered, m_nodes, &FlowScene::load);
 }
 
 MainWindow::~MainWindow()

### 2018-03-13 include/RenderableManager.h
added ribbons geo test shader

@@ -78,7 +78,7 @@ class RenderableManager
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief our shaders
 		//----------------------------------------------------------------------------------------------------------------------
-		Shader meshShader, curvesShader;
+		Shader meshShader, curvesShader, curvesRibbonShader;
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief a pointer to our viewport
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-13 src/RenderableManager.cpp

@@ -19,6 +19,7 @@ RenderableManager::RenderableManager()
 {
 	meshShader = Shader("shaders/render/meshShader_vert.glsl", "shaders/render/meshShader_frag.glsl");
 	curvesShader = Shader("shaders/render/curvesShader_vert.glsl", "shaders/render/curvesShader_frag.glsl", "shaders/render/curvesShader_geo.glsl");
+	curvesRibbonShader = Shader("shaders/render/curvesShader_vert.glsl", "shaders/render/curvesShader_frag.glsl", "shaders/render/curvesRibbonShader_geo.glsl");
 }
 
 //----------------------------------------------------------------------------------------------------------------------
@@ -34,6 +35,7 @@ void RenderableManager::cleanUpAll()
 {
 	meshShader.cleanUp();
 	curvesShader.cleanUp();
+	curvesRibbonShader.cleanUp();
 
 	for (auto renderable : m_renderables)
 	{
@@ -47,8 +49,10 @@ void RenderableManager::recompileShaders()
 {
 	meshShader.cleanUp();
 	curvesShader.cleanUp();
+	curvesRibbonShader.cleanUp();
 	meshShader = Shader("shaders/render/meshShader_vert.glsl", "shaders/render/meshShader_frag.glsl");
 	curvesShader = Shader("shaders/render/curvesShader_vert.glsl", "shaders/render/curvesShader_frag.glsl", "shaders/render/curvesShader_geo.glsl");
+	curvesRibbonShader = Shader("shaders/render/curvesShader_vert.glsl", "shaders/render/curvesShader_frag.glsl", "shaders/render/curvesRibbonShader_geo.glsl");
 }
 
 //----------------------------------------------------------------------------------------------------------------------
@@ -125,11 +129,13 @@ void RenderableManager::drawRenderables()
 
 				case CURVES:
 					// std::cout<<"CURVES draw call\n";
-					curvesShader.use();
-					loadMatricesToShader(curvesShader.getID());
+					// curvesShader.use();
+					curvesRibbonShader.use();
+					// loadMatricesToShader(curvesShader.getID());
+					loadMatricesToShader(curvesRibbonShader.getID());
 					// auto r = std::static_pointer_cast<CurvesRenderable>(renderable);
 					// col = std::static_pointer_cast<CurvesRenderable>(renderable)->getColour();
[…]

### 2018-03-13 include/CurvesRenderable.h
everything seems to work

@@ -24,7 +24,7 @@ class CurvesRenderable : public Renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief get the type of this renderable
 		//----------------------------------------------------------------------------------------------------------------------
-		renderableType getType() const override {return CURVES;}
+		RenderableType getType() const override {return m_curveRenderType;}
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief generate the SSBO for the renderable
 		//----------------------------------------------------------------------------------------------------------------------
@@ -38,9 +38,14 @@ class CurvesRenderable : public Renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		void setSourceSSBO(GLuint _SSBO);
 		//----------------------------------------------------------------------------------------------------------------------
-		/// @brief set the colour of this renderable
+		/// @brief set the compute type of this renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		void setCurveType(CurveType _type);
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief set the render type of this renderable
+		//----------------------------------------------------------------------------------------------------------------------
+		inline void setRibbons() {m_curveRenderType = CURVES_RIBBONS;}
+		inline void setLines() {m_curveRenderType = CURVES_LINES;}
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief set the mesh
 		//----------------------------------------------------------------------------------------------------------------------
@@ -50,6 +55,10 @@ class CurvesRenderable : public Renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		void setColour(glm::vec3 const &_col);
 		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief get the width for the curve ribbons
+		//----------------------------------------------------------------------------------------------------------------------
+		void setWidth(float _width) {m_ribbonWidth = _width;}
+		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief get the colour of this renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		glm::vec3 getColour() const {return m_colour;}
@@ -57,12 +66,20 @@ class CurvesRenderable : public Renderable
 		/// @brief get the number of indices in this renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		unsigned int getIndices() const {return m_indices;}
+		//----------------------------------------------------------------------------------------------------------------------
[…]

### 2018-03-13 include/CurvesRendererModel.h

@@ -3,7 +3,7 @@
 
 #include "CurvesRenderable.h"
 
-#include "widgets/MeshRendererModelWidget.h"
+#include "widgets/CurvesRendererModelWidget.h"
 
 #include <iostream>
 
@@ -65,6 +65,14 @@ class CurvesRendererModel : public NodeDataModel
 		/// @brief set the visibility of our renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		void setVisibility();
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief set the render type of our renderable
+		//----------------------------------------------------------------------------------------------------------------------
+		void setRenderType();
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief set the ribbon width of our renderable
+		//----------------------------------------------------------------------------------------------------------------------
+		void setWidth();
 
 	protected:
 		//----------------------------------------------------------------------------------------------------------------------
@@ -85,7 +93,7 @@ class CurvesRendererModel : public NodeDataModel
 		/// @brief m_embedded, m_ui our qt ui widget and contents from the ui file
 		//----------------------------------------------------------------------------------------------------------------------
 		QWidget* m_embedded;
-		Ui::MeshRendererModelWidget* m_ui;
+		Ui::CurvesRendererModelWidget* m_ui;
 };
 
 #endif // CURVESRENDERERMODEL_H

### 2018-03-13 include/MeshRenderable.h

@@ -24,7 +24,7 @@ class MeshRenderable : public Renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief get the type of this renderable
 		//----------------------------------------------------------------------------------------------------------------------
-		renderableType getType() const override {return MESH;}
+		RenderableType getType() const override {return MESH;}
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief generate the VAO for the renderable
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-13 include/Renderable.h

@@ -3,10 +3,11 @@
 
 #include <iostream>
 
-enum renderableType
+enum RenderableType
 {
 	DEFAULT,
-	CURVES,
+	CURVES_LINES,
+	CURVES_RIBBONS,
 	MESH
 };
 
@@ -28,7 +29,7 @@ class Renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief get the type of this renderable
 		//----------------------------------------------------------------------------------------------------------------------
-		virtual renderableType getType() const {return DEFAULT;}
+		virtual RenderableType getType() const {return DEFAULT;}
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief draw this renderable
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-13 include/widgets/CurvesRendererModelWidget.h

@@ -1,0 +1,149 @@
+/********************************************************************************
+** Form generated from reading UI file 'CurvesRendererModelWidget.ui'
+**
+** Created by: Qt User Interface Compiler version 5.9.2
+**
+** WARNING! All changes made in this file will be lost when recompiling UI file!
+********************************************************************************/
+
+#ifndef CURVESRENDERERMODELWIDGET_H
+#define CURVESRENDERERMODELWIDGET_H
+
+#include <QtCore/QVariant>
+#include <QtWidgets/QAction>
+#include <QtWidgets/QApplication>
+#include <QtWidgets/QButtonGroup>
+#include <QtWidgets/QCheckBox>
+#include <QtWidgets/QDoubleSpinBox>
+#include <QtWidgets/QFormLayout>
+#include <QtWidgets/QHeaderView>
+#include <QtWidgets/QLabel>
+#include <QtWidgets/QWidget>
+
+QT_BEGIN_NAMESPACE
+
+class Ui_CurvesRendererModelWidget
+{
+public:
+    QFormLayout *formLayout;
+    QLabel *redAmountLabel;
+    QDoubleSpinBox *redAmountDoubleSpinBox;
+    QLabel *greenAmountLabel;
+    QDoubleSpinBox *greenAmountDoubleSpinBox;
+    QLabel *blueAmountLabel;
+    QDoubleSpinBox *blueAmountDoubleSpinBox;
+    QLabel *visibilityLabel;
+    QCheckBox *visibilityCheckBox;
+    QLabel *ribbonsLabel;
+    QCheckBox *ribbonsCheckBox;
+    QLabel *widthLabel;
+    QDoubleSpinBox *widthDoubleSpinBox;
+
+    void setupUi(QWidget *CurvesRendererModelWidget)
+    {
+        if (CurvesRendererModelWidget->objectName().isEmpty())
+            CurvesRendererModelWidget->setObjectName(QStringLiteral("CurvesRendererModelWidget"));
+        CurvesRendererModelWidget->resize(136, 153);
+        formLayout = new QFormLayout(CurvesRendererModelWidget);
+        formLayout->setObjectName(QStringLiteral("formLayout"));
+        redAmountLabel = new QLabel(CurvesRendererModelWidget);
+        redAmountLabel->setObjectName(QStringLiteral("redAmountLabel"));
+
+        formLayout->setWidget(0, QFormLayout::LabelRole, redAmountLabel);
+
[…]

### 2018-03-13 src/CurvesRenderable.cpp

@@ -5,11 +5,13 @@
 CurvesRenderable::CurvesRenderable()
 {
 	std::cout<<"CurvesRenderable()\n";
+	m_curveRenderType = CURVES_RIBBONS;
 	m_emptyVAO = 0;
 	m_SSBO = 0;
 	m_curves = Curves();
 	m_colour = glm::vec3(1.0, 0.0, 0.0);
 	m_indices = 0;
+	m_ribbonWidth = 0.005;
 }
 
 //----------------------------------------------------------------------------------------------------------------------

### 2018-03-13 src/CurvesRendererModel.cpp

@@ -8,7 +8,7 @@
 
 CurvesRendererModel::CurvesRendererModel() :
 	m_embedded(new QWidget()),
-	m_ui(new Ui::MeshRendererModelWidget())
+	m_ui(new Ui::CurvesRendererModelWidget())
 {
 	// setup ui
 	m_ui->setupUi(m_embedded);
@@ -17,7 +17,9 @@ CurvesRendererModel::CurvesRendererModel() :
 	connect(m_ui->redAmountDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setColour()));
 	connect(m_ui->greenAmountDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setColour()));
 	connect(m_ui->blueAmountDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setColour()));
+	connect(m_ui->widthDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setWidth()));
 	connect(m_ui->visibilityCheckBox, SIGNAL(stateChanged(int)), this, SLOT(setVisibility()));
+	connect(m_ui->ribbonsCheckBox, SIGNAL(stateChanged(int)), this, SLOT(setRenderType()));
 	m_curvesRenderable = std::make_shared<CurvesRenderable>();
 }
 
@@ -102,5 +104,24 @@ void CurvesRendererModel::setColour()
 void CurvesRendererModel::setVisibility()
 {
 	m_curvesRenderable->setVisibility(m_ui->visibilityCheckBox->isChecked());
+	RenderableManager::getInstance()->updateViewport();
+}
+
+void CurvesRendererModel::setRenderType()
+{
+	if (m_ui->ribbonsCheckBox->isChecked())
+	{
+		m_curvesRenderable->setRibbons();
+	}
+	else
+	{
+		m_curvesRenderable->setLines();
+	}
+	RenderableManager::getInstance()->updateViewport();
+}
+
+void CurvesRendererModel::setWidth()
+{
+	m_curvesRenderable->setWidth(m_ui->widthDoubleSpinBox->value());
 	RenderableManager::getInstance()->updateViewport();
 }

### 2018-03-13 src/RenderableManager.cpp

@@ -127,16 +127,20 @@ void RenderableManager::drawRenderables()
 					renderable->draw();
 					break;
 
-				case CURVES:
-					// std::cout<<"CURVES draw call\n";
-					// curvesShader.use();
-					curvesRibbonShader.use();
-					// loadMatricesToShader(curvesShader.getID());
-					loadMatricesToShader(curvesRibbonShader.getID());
+				case CURVES_LINES:
+					curvesShader.use();
+					loadMatricesToShader(curvesShader.getID());
 					// auto r = std::static_pointer_cast<CurvesRenderable>(renderable);
 					// col = std::static_pointer_cast<CurvesRenderable>(renderable)->getColour();
+					glUniform3fv(glGetUniformLocation(curvesShader.getID(), "u_colour"), 1, glm::value_ptr(std::static_pointer_cast<CurvesRenderable>(renderable)->getColour()));
+					renderable->draw();
+					break;
+
+				case CURVES_RIBBONS:
+					curvesRibbonShader.use();
+					loadMatricesToShader(curvesRibbonShader.getID());
 					glUniform3fv(glGetUniformLocation(curvesRibbonShader.getID(), "u_colour"), 1, glm::value_ptr(std::static_pointer_cast<CurvesRenderable>(renderable)->getColour()));
-					// glUniform1ui(glGetUniformLocation(curvesShader.getID(), "u_numIndices"), std::static_pointer_cast<CurvesRenderable>(renderable)->getIndices());
+					glUniform1f(glGetUniformLocation(curvesRibbonShader.getID(), "u_width"), std::static_pointer_cast<CurvesRenderable>(renderable)->getWidth());
 					renderable->draw();
 					break;
 

### 2018-03-13 include/OrbitingCamera.h
added camera reset & shading of fur

@@ -31,6 +31,8 @@ class OrbitingCamera
 		void processMouseMovement(float xoffset, float yoffset, bool constrainPitch = true) noexcept;
 		//----------------------------------------------------------------------------------------------------------------------
 		void processMouseScroll(float _yoffset) noexcept;
+		//----------------------------------------------------------------------------------------------------------------------
+		void resetPos() noexcept;
 
 	private :
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-13 src/CurvesRendererModel.cpp

@@ -110,13 +110,9 @@ void CurvesRendererModel::setVisibility()
 void CurvesRendererModel::setRenderType()
 {
 	if (m_ui->ribbonsCheckBox->isChecked())
-	{
 		m_curvesRenderable->setRibbons();
-	}
 	else
-	{
 		m_curvesRenderable->setLines();
-	}
 	RenderableManager::getInstance()->updateViewport();
 }
 

### 2018-03-13 src/OrbitingCamera.cpp

@@ -76,3 +76,8 @@ void OrbitingCamera::calculateOrbitPos() noexcept
 	m_eye.y = m_radius * sin(m_phi);
 	m_eye.z = m_radius * cos(m_theta) * cos(m_phi);
 }
+
+void OrbitingCamera::resetPos() noexcept
+{
+	m_look = glm::vec3(0.0);
+}

### 2018-03-13 src/Viewport.cpp

@@ -133,6 +133,8 @@ void Viewport::keyPressEvent(QKeyEvent *_event)
 		case Qt::Key_S : glPolygonMode(GL_FRONT_AND_BACK,GL_FILL); break;
 		// show full screen
 		case Qt::Key_F : showFullScreen(); break;
+		// center camera
+		case Qt::Key_C : m_cam.resetPos(); break;
 		// show windowed
 		case Qt::Key_N : showNormal(); break;
 		default : break;

### 2018-03-13 include/CurvesRenderable.h
added shading controls

@@ -70,6 +70,22 @@ class CurvesRenderable : public Renderable
 		/// @brief get the width for the curve ribbons
 		//----------------------------------------------------------------------------------------------------------------------
 		float getWidth() const {return m_ribbonWidth;}
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief set the Roughness of this renderable
+		//----------------------------------------------------------------------------------------------------------------------
+		void setRoughness(float const &_roughness);
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief get the Roughness of this renderable
+		//----------------------------------------------------------------------------------------------------------------------
+		float getRoughness() const {return m_roughness;}
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief set the Metallic of this renderable
+		//----------------------------------------------------------------------------------------------------------------------
+		void setMetallic(float const &_metallic);
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief get the Metallic of this renderable
+		//----------------------------------------------------------------------------------------------------------------------
+		float getMetallic() const {return m_metallic;}
 
 	private:
 		//----------------------------------------------------------------------------------------------------------------------
@@ -108,6 +124,10 @@ class CurvesRenderable : public Renderable
 		/// @brief m_ribbonWidth the width of our renderable ribbons
 		//----------------------------------------------------------------------------------------------------------------------
 		float m_ribbonWidth;
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief m_roughness, m_metallic roughness and metallic
+		//----------------------------------------------------------------------------------------------------------------------
+		float m_roughness, m_metallic;
 };
 
 #endif // CURVESRENDERABLE_H

### 2018-03-13 include/CurvesRendererModel.h

@@ -62,6 +62,10 @@ class CurvesRendererModel : public NodeDataModel
 		//----------------------------------------------------------------------------------------------------------------------
 		void setColour();
 		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief set uniforms
+		//----------------------------------------------------------------------------------------------------------------------
+		void setMetallicRoughness();
+		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief set the visibility of our renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		void setVisibility();

### 2018-03-13 include/MeshRenderable.h

@@ -45,6 +45,23 @@ class MeshRenderable : public Renderable
 		/// @brief get the colour of this renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		glm::vec3 getColour() const {return m_colour;}
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief set the Roughness of this renderable
+		//----------------------------------------------------------------------------------------------------------------------
+		void setRoughness(float const &_roughness);
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief get the Roughness of this renderable
+		//----------------------------------------------------------------------------------------------------------------------
+		float getRoughness() const {return m_roughness;}
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief set the Metallic of this renderable
+		//----------------------------------------------------------------------------------------------------------------------
+		void setMetallic(float const &_metallic);
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief get the Metallic of this renderable
+		//----------------------------------------------------------------------------------------------------------------------
+		float getMetallic() const {return m_metallic;}
+
 
 	private:
 		//----------------------------------------------------------------------------------------------------------------------
@@ -63,6 +80,10 @@ class MeshRenderable : public Renderable
 		/// @brief m_colour the colour of our renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		glm::vec3 m_colour;
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief m_roughness, m_metallic roughness and metallic
+		//----------------------------------------------------------------------------------------------------------------------
+		float m_roughness, m_metallic;
 };
 
 #endif // MESHRENDERABLE_H

### 2018-03-13 include/MeshRendererModel.h

@@ -62,6 +62,10 @@ class MeshRendererModel : public NodeDataModel
 		//----------------------------------------------------------------------------------------------------------------------
 		void setColour();
 		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief set uniforms
+		//----------------------------------------------------------------------------------------------------------------------
+		void setMetallicRoughness();
+		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief set the visibility of our renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		void setVisibility();

### 2018-03-13 include/RenderableManager.h

@@ -74,7 +74,7 @@ class RenderableManager
 		//----------------------------------------------------------------------------------------------------------------------
 		void updateViewport() {m_viewport->update();}
 
-	private:
+	// private:
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief our shaders
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-13 include/widgets/CurvesRendererModelWidget.h

@@ -1,7 +1,7 @@
 /********************************************************************************
 ** Form generated from reading UI file 'CurvesRendererModelWidget.ui'
 **
-** Created by: Qt User Interface Compiler version 5.9.2
+** Created by: Qt User Interface Compiler version 5.9.0
 **
 ** WARNING! All changes made in this file will be lost when recompiling UI file!
 ********************************************************************************/
@@ -38,12 +38,16 @@ public:
     QCheckBox *ribbonsCheckBox;
     QLabel *widthLabel;
     QDoubleSpinBox *widthDoubleSpinBox;
+    QLabel *roughnessLabel;
+    QLabel *metallicLabel;
+    QDoubleSpinBox *roughnessDoubleSpinBox;
+    QDoubleSpinBox *metallicDoubleSpinBox;
 
     void setupUi(QWidget *CurvesRendererModelWidget)
     {
         if (CurvesRendererModelWidget->objectName().isEmpty())
             CurvesRendererModelWidget->setObjectName(QStringLiteral("CurvesRendererModelWidget"));
-        CurvesRendererModelWidget->resize(136, 153);
+        CurvesRendererModelWidget->resize(152, 258);
         formLayout = new QFormLayout(CurvesRendererModelWidget);
         formLayout->setObjectName(QStringLiteral("formLayout"));
         redAmountLabel = new QLabel(CurvesRendererModelWidget);
@@ -86,29 +90,29 @@ public:
         visibilityLabel = new QLabel(CurvesRendererModelWidget);
         visibilityLabel->setObjectName(QStringLiteral("visibilityLabel"));
 
-        formLayout->setWidget(3, QFormLayout::LabelRole, visibilityLabel);
+        formLayout->setWidget(5, QFormLayout::LabelRole, visibilityLabel);
 
         visibilityCheckBox = new QCheckBox(CurvesRendererModelWidget);
         visibilityCheckBox->setObjectName(QStringLiteral("visibilityCheckBox"));
         visibilityCheckBox->setChecked(true);
 
-        formLayout->setWidget(3, QFormLayout::FieldRole, visibilityCheckBox);
+        formLayout->setWidget(5, QFormLayout::FieldRole, visibilityCheckBox);
 
         ribbonsLabel = new QLabel(CurvesRendererModelWidget);
[…]

### 2018-03-13 include/widgets/MeshRendererModelWidget.h

@@ -1,7 +1,7 @@
 /********************************************************************************
 ** Form generated from reading UI file 'MeshRendererModelWidget.ui'
 **
-** Created by: Qt User Interface Compiler version 5.9.2
+** Created by: Qt User Interface Compiler version 5.9.0
 **
 ** WARNING! All changes made in this file will be lost when recompiling UI file!
 ********************************************************************************/
@@ -34,12 +34,16 @@ public:
     QDoubleSpinBox *blueAmountDoubleSpinBox;
     QLabel *visibilityLabel;
     QCheckBox *visibilityCheckBox;
+    QLabel *roughnessLabel;
+    QLabel *metallicLabel;
+    QDoubleSpinBox *roughnessDoubleSpinBox;
+    QDoubleSpinBox *metallicDoubleSpinBox;
 
     void setupUi(QWidget *MeshRendererModelWidget)
     {
         if (MeshRendererModelWidget->objectName().isEmpty())
             MeshRendererModelWidget->setObjectName(QStringLiteral("MeshRendererModelWidget"));
-        MeshRendererModelWidget->resize(129, 108);
+        MeshRendererModelWidget->resize(145, 201);
         formLayout = new QFormLayout(MeshRendererModelWidget);
         formLayout->setObjectName(QStringLiteral("formLayout"));
         redAmountLabel = new QLabel(MeshRendererModelWidget);
@@ -82,13 +86,39 @@ public:
         visibilityLabel = new QLabel(MeshRendererModelWidget);
         visibilityLabel->setObjectName(QStringLiteral("visibilityLabel"));
 
-        formLayout->setWidget(3, QFormLayout::LabelRole, visibilityLabel);
+        formLayout->setWidget(5, QFormLayout::LabelRole, visibilityLabel);
 
         visibilityCheckBox = new QCheckBox(MeshRendererModelWidget);
         visibilityCheckBox->setObjectName(QStringLiteral("visibilityCheckBox"));
         visibilityCheckBox->setChecked(true);
 
-        formLayout->setWidget(3, QFormLayout::FieldRole, visibilityCheckBox);
+        formLayout->setWidget(5, QFormLayout::FieldRole, visibilityCheckBox);
+
+        roughnessLabel = new QLabel(MeshRendererModelWidget);
+        roughnessLabel->setObjectName(QStringLiteral("roughnessLabel"));
[…]

### 2018-03-13 src/CurvesRenderable.cpp

@@ -12,6 +12,8 @@ CurvesRenderable::CurvesRenderable()
 	m_colour = glm::vec3(1.0, 0.0, 0.0);
 	m_indices = 0;
 	m_ribbonWidth = 0.005;
+	m_roughness = 0.5;
+	m_metallic = 0.5;
 }
 
 //----------------------------------------------------------------------------------------------------------------------
@@ -118,4 +120,18 @@ void CurvesRenderable::setCurves(Curves const &_curves)
 void CurvesRenderable::setColour(glm::vec3 const &_col)
 {
 	m_colour = _col;
+}
+
+//----------------------------------------------------------------------------------------------------------------------
+
+void CurvesRenderable::setMetallic(float const &_metallic)
+{
+	m_metallic = _metallic;
+}
+
+//----------------------------------------------------------------------------------------------------------------------
+
+void CurvesRenderable::setRoughness(float const &_roughness)
+{
+	m_roughness = _roughness;
 }

### 2018-03-13 src/CurvesRendererModel.cpp

@@ -18,8 +18,10 @@ CurvesRendererModel::CurvesRendererModel() :
 	connect(m_ui->greenAmountDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setColour()));
 	connect(m_ui->blueAmountDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setColour()));
 	connect(m_ui->widthDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setWidth()));
-	connect(m_ui->visibilityCheckBox, SIGNAL(stateChanged(int)), this, SLOT(setVisibility()));
+	connect(m_ui->roughnessDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setMetallicRoughness()));
+	connect(m_ui->metallicDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setMetallicRoughness()));
 	connect(m_ui->ribbonsCheckBox, SIGNAL(stateChanged(int)), this, SLOT(setRenderType()));
+	connect(m_ui->visibilityCheckBox, SIGNAL(stateChanged(int)), this, SLOT(setVisibility()));
 	m_curvesRenderable = std::make_shared<CurvesRenderable>();
 }
 
@@ -119,5 +121,12 @@ void CurvesRendererModel::setRenderType()
 void CurvesRendererModel::setWidth()
 {
 	m_curvesRenderable->setWidth(m_ui->widthDoubleSpinBox->value());
+	RenderableManager::getInstance()->updateViewport();
+}
+
+void CurvesRendererModel::setMetallicRoughness()
+{
+	m_curvesRenderable->setRoughness(m_ui->roughnessDoubleSpinBox->value());
+	m_curvesRenderable->setMetallic(m_ui->metallicDoubleSpinBox->value());
 	RenderableManager::getInstance()->updateViewport();
 }

### 2018-03-13 src/MeshRenderable.cpp

@@ -8,6 +8,8 @@ MeshRenderable::MeshRenderable()
 	m_indices = 0;
 	m_mesh = Mesh();
 	m_colour = glm::vec3(1.0, 0.0, 0.0);
+	m_roughness = 0.5;
+	m_metallic = 0.5;
 }
 
 //----------------------------------------------------------------------------------------------------------------------
@@ -107,4 +109,18 @@ void MeshRenderable::setMesh(Mesh const &_mesh)
 void MeshRenderable::setColour(glm::vec3 const &_col)
 {
 	m_colour = _col;
+}
+
+//----------------------------------------------------------------------------------------------------------------------
+
+void MeshRenderable::setMetallic(float const &_metallic)
+{
+	m_metallic = _metallic;
+}
+
+//----------------------------------------------------------------------------------------------------------------------
+
+void MeshRenderable::setRoughness(float const &_roughness)
+{
+	m_roughness = _roughness;
 }

### 2018-03-13 src/MeshRendererModel.cpp

@@ -17,6 +17,8 @@ MeshRendererModel::MeshRendererModel() :
 	connect(m_ui->redAmountDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setColour()));
 	connect(m_ui->greenAmountDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setColour()));
 	connect(m_ui->blueAmountDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setColour()));
+	connect(m_ui->roughnessDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setMetallicRoughness()));
+	connect(m_ui->metallicDoubleSpinBox, SIGNAL(valueChanged(double)), this, SLOT(setMetallicRoughness()));
 	connect(m_ui->visibilityCheckBox, SIGNAL(stateChanged(int)), this, SLOT(setVisibility()));
 	m_meshRenderable = std::make_shared<MeshRenderable>();
 }
@@ -100,5 +102,12 @@ void MeshRendererModel::setColour()
 void MeshRendererModel::setVisibility()
 {
 	m_meshRenderable->setVisibility(m_ui->visibilityCheckBox->isChecked());
+	RenderableManager::getInstance()->updateViewport();
+}
+
+void MeshRendererModel::setMetallicRoughness()
+{
+	m_meshRenderable->setRoughness(m_ui->roughnessDoubleSpinBox->value());
+	m_meshRenderable->setMetallic(m_ui->metallicDoubleSpinBox->value());
 	RenderableManager::getInstance()->updateViewport();
 }

### 2018-03-13 src/RenderableManager.cpp

@@ -124,6 +124,8 @@ void RenderableManager::drawRenderables()
 					loadMatricesToShader(meshShader.getID());
 					col = std::static_pointer_cast<MeshRenderable>(renderable)->getColour();
 					glUniform3fv(glGetUniformLocation(meshShader.getID(), "u_colour"), 1, glm::value_ptr(col));
+					glUniform1f(glGetUniformLocation(meshShader.getID(), "u_roughness"), std::static_pointer_cast<MeshRenderable>(renderable)->getRoughness());
+					glUniform1f(glGetUniformLocation(meshShader.getID(), "u_metallic"), std::static_pointer_cast<MeshRenderable>(renderable)->getMetallic());
 					renderable->draw();
 					break;
 
@@ -133,6 +135,8 @@ void RenderableManager::drawRenderables()
 					// auto r = std::static_pointer_cast<CurvesRenderable>(renderable);
 					// col = std::static_pointer_cast<CurvesRenderable>(renderable)->getColour();
 					glUniform3fv(glGetUniformLocation(curvesShader.getID(), "u_colour"), 1, glm::value_ptr(std::static_pointer_cast<CurvesRenderable>(renderable)->getColour()));
+					glUniform1f(glGetUniformLocation(curvesShader.getID(), "u_roughness"), std::static_pointer_cast<CurvesRenderable>(renderable)->getRoughness());
+					glUniform1f(glGetUniformLocation(curvesShader.getID(), "u_metallic"), std::static_pointer_cast<CurvesRenderable>(renderable)->getMetallic());
 					renderable->draw();
 					break;
 
@@ -141,6 +145,8 @@ void RenderableManager::drawRenderables()
 					loadMatricesToShader(curvesRibbonShader.getID());
 					glUniform3fv(glGetUniformLocation(curvesRibbonShader.getID(), "u_colour"), 1, glm::value_ptr(std::static_pointer_cast<CurvesRenderable>(renderable)->getColour()));
 					glUniform1f(glGetUniformLocation(curvesRibbonShader.getID(), "u_width"), std::static_pointer_cast<CurvesRenderable>(renderable)->getWidth());
+					glUniform1f(glGetUniformLocation(curvesRibbonShader.getID(), "u_roughness"), std::static_pointer_cast<CurvesRenderable>(renderable)->getRoughness());
[…]

### 2018-03-14 docs/paper.tex
more on docs

@@
 For this project I developed a procedural fur system, with the aim of exploring the feasibility of offloading computation onto the GPU within artist tools.
-The final artefact is an application which serves to interface with the fur system API that I have developed.
+The final artefact is an application which serves to interface with the fur system API which I have developed.
 This report primarily documents the implementation of the API, as well as my findings.
@@
 Whilst many larger companies opt to develop their own software for handling fur, such as MPC's Furtility \cite{furtility}, there are a number of commercially available solutions that are also widely used in production.
+
 The most popular of these is XGen, originally developed by Walt Disney Animation Studios, which has been included in Autodesk's Maya since 2014.
@@
 
+\subsection{GPU Acceleration} \label{sec:existingGPU}
+Whilst it is not clear to what extent each of these solutions make use of GPU acceleration, they all perform fast enough to allow the user to manipulate them interactively.
+I therefore decided to monitor the CPU and GPU usage using \verb|htop| and \verb|nvidia-smi| respectively, before and after adding XGen content to a simple maya scene, to determine where the majority of data was being computed.
+
+After applying the default XGen description onto a simple sphere with 760 triangles, there was no noticeable increase in both GPU utilisation and GPU memory usage.
+Increasing the density of the primitives from 1.0 to 10000.0 saw an increase of VRAM usage of roughly 350MiB, which I would attribute to storing the primitives in a buffer on the GPU, but there was no discernable increase in GPU utilisation.
+Increasing the density further to 10,000 saw a noticable spike in CPU usage, hitting 100 percent across two threads, leading me to believe that the primitive distribution computation is being done on the CPU.
+
+However, when applying modifiers to the XGen description I noticed a small increase of roughly 2MiB in VRAM usage for the first modifier, yet no further increases for modifiers added after that.
+As this increase is quite small I would attribute it to storing the modifier attributes on the GPU, though I still noticed spikes in CPU usage when recalculating the XGen description once the modifier was added, which makes it inconclusive as to whether XGen uses any kind of GPU acceleration.
+
+% on maya load: CPU up to 8 percent when moving mouse GPU very low but 40 percent when moving viewport 42MiB
+% before xgen on sphere 36mib
+% xgen on sphere density 1.0 36mib GPU
+% xgen on sphere density 100.0 40mib GPU
+% xgen CPU spiked up to 100 on a single thread when changing density
+% xgen on sphere density 1000.0 56mib GPU
+% xgen noise on sphere density 1000.0 63mib GPU, CPU spiked but only to 20%
+% xgen on sphere density to 10,000 caused spike on 2 threads
+
 \section{Implementation} \label{sec:implementation}
-Method Text.
+From my research it was clear that there is an apparent lack of standalone software for authoring fur; it makes sense that existing fur software are developed as software plugins, allowing the tool to make use of their host DCCs existing functionality.
+
+However, by developing the system as a simple API one could make it DCC agnostic, which is why I chose to develop my system as a pairing of a simple API with a standalone tool which interfaces with it.
+By developing a simple API it could allow for future implementations as plugins within DCCs, providing the data exchange formats are compatible with the host DCC.
 
@@
 
-write about obj loading
-
-write about distributing https://stackoverflow.com/questions/9294316/distribute-points-on-mesh-according-to-density
+write about distributing https://stackoverflow.com/questions/9294316/
+distribute-points-on-mesh-according-to-density
 
 write about GPU
+
+\subsubsection{Mesh Loading} \label{sec:objLoading}
+The first problem I tackled was that of loading mesh geometry into the system, and storing it in a Shader Storage Buffer Object (SSBO) for later use in a compute shader.
+I opted to implement a Wavefront OBJ loader as they are (deceptively) simple to parse, consisting of a list of vertex positions, vertex normals, vertex texture coordinates, and faces defined as a list of indices.
+
+The most common way of storing OBJ data in OpenGL buffers, as described in \cite{objLoading}, is to use a \verb|GL_ARRAY_BUFFER| object to store the vertex data and a \verb|GL_ELEMENT_ARRAY_BUFFER| to store the indices of the \verb|GL_ARRAY_BUFFER| that make up each face.
+However for ease of use in compute shaders I decided instead to store each face in a \verb|GL_SHADER_STORAGE_BUFFER|, where each face consists of three sets of positions, normals, and texture coordinates.
+Whilst this is less efficient, I found that the overhead in GPU memory usage was negligible, and that this method allowed for much simpler usage in a compute shader.
 

### 2018-03-15 docs/paper.tex
more on docs

@@
 
-write about distributing https://stackoverflow.com/questions/9294316/
-distribute-points-on-mesh-according-to-density
+In this section I will explain the development of each of the major components, not necessarily in chronological order, but in the order regarding to how their data is passed through the system.
 
-write about GPU
+% write about distributing https://stackoverflow.com/questions/9294316/
+% distribute-points-on-mesh-according-to-density
+
+% write about GPU
 
 \subsubsection{Mesh Loading} \label{sec:objLoading}
-The first problem I tackled was that of loading mesh geometry into the system, and storing it in a Shader Storage Buffer Object (SSBO) for later use in a compute shader.
+% The first problem I tackled was that of loading mesh geometry into the system, and storing it in a Shader Storage Buffer Object (SSBO) for later use in a compute shader.
 I opted to implement a Wavefront OBJ loader as they are (deceptively) simple to parse, consisting of a list of vertex positions, vertex normals, vertex texture coordinates, and faces defined as a list of indices.
+I created an OBJ loader node using the NodeEditor library, which parses
+
+The first problem I tackled was that of loading mesh geometry into the system, in such a way that would allow for later computation on both the GPU and CPU.
+I opted to implement a Wavefront OBJ loader as they are (deceptively) simple to parse, consisting of a list of vertex positions, vertex normals, vertex texture coordinates, and faces as defined by a list of indices.
+This data structure formed the basis of my 'Mesh' class.
+Using the NodeEditor library I then created a node which loads OBJ files, and outputs the 'Mesh' object through it's output data port.
+
+\subsubsection{Curve Distributors} \label{sec:distributors}
+
+In order to generate curves for the system to manipulate, it was necessary to create a node that handles the distribution of curves onto the surface of a mesh.
+This node needs to take 'Mesh' object data as input, and output 'Curves' object data.
+A 'Curves' object was implemented which consists an array of 'Curve' objects, which in turn consist of an array 5 vertex positions that construct the curve.
+Distributor nodes would also need to alternate between performing the distribution process on the CPU or on the GPU, which would in turn affect whether the node outputs a 'Curves' object, or an ID for where the curves data is stored on the GPU.
+
+With these parameters in mind I decided to implement an abstract distributor node class, which contains a virtual function for performing the distribution, and handles the callbacks for when input data is changed and the distribution needs to be updated.
+This makes the system somewhat extensible, allowing for different distribution patterns to be implemented with ease.
+The node's data type was also extented to be able to store either 'Curves' objects from the CPU, or an ID for an OpenGL buffer object, with an enumerated value to identify which of these is currently in use.
+This not only allows the distributor to alternate between CPU and GPU computation arbitrarily, but also allows nodes recieving this data to process it accordingly.
 

### 2018-03-16 docs/paper.tex
more on docs

@@
 
+There are many different methods one could use to distribute points on an abritrary mesh, however for replicating creature fur it is desirable to choose a method that distributes primitives at random positions across the surface, yet roughly evenly spaced from neighboring primitives.
+A popular method used to achieve this kind of distribution is referred to as 'Dart Throwing', which relies on checking the distance to neighboring samples after a new sample has been generated, and discarding the sample if the samples are too close.
+Whilst these kinds of methods produce good results, they can be inefficient as in certain situations many samples may be generated and rejected, before a sample is accepted.
+An optimisation for this is described in \textit{Dart Throwing on Surfaces} \cite{dartThrowing}, which relies on removing areas of the mesh that are already covered by a sample, by means of triangle subdivision.
+
+Unfortunately, parallelizing dart thowing methods is not trivial, making them unsuitable for my desired usage of OpenGL compute shaders.
+Whilst sampling methods suitable for parallel processing do exist, such as those described in \textit{Parallel Poisson Disk Sampling with Spectrum Analysis on Surfaces} \cite{parallelPoisson}, I instead opted to implement a naive random samling method as it was much easier to parallelize.
+
+\begin{enumerate}
+\item Generate a uniform random number in the range [0.0 - 1.0].
+\end{enumerate}
+
 The most common way of storing OBJ data in OpenGL buffers, as described in \cite{objLoading}, is to use a \verb|GL_ARRAY_BUFFER| object to store the vertex data and a \verb|GL_ELEMENT_ARRAY_BUFFER| to store the indices of the \verb|GL_ARRAY_BUFFER| that make up each face.

### 2018-03-16 include/ComputeShaderManager.h
gathering metrics

@@ -7,6 +7,8 @@
 
 #include <GL/glew.h>
 
+#include <chrono>
+
 class ComputeShaderManager
 {
 	private:
@@ -67,11 +69,25 @@ class ComputeShaderManager
 		//----------------------------------------------------------------------------------------------------------------------
 		void clumpCurvesOperator(unsigned int curvesSSBO, unsigned int clumpCurvesSSBO, float envelope, bool preserveLength);
 
+
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief start chrono timer
+		//----------------------------------------------------------------------------------------------------------------------
+		void startTimer();
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief end timer and print elapsed
+		//----------------------------------------------------------------------------------------------------------------------
+		void endTimer();
+
 	private:
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief our compute shaders
 		//----------------------------------------------------------------------------------------------------------------------
 		Shader randomDistributionShader, bendCurvesShader, noiseCurvesShader, clumpCurvesShader;
+		//----------------------------------------------------------------------------------------------------------------------
+		/// @brief our timer for measuring performance
+		//----------------------------------------------------------------------------------------------------------------------
+		std::chrono::system_clock::time_point m_timer;
 };
 
 #endif // COMPUTESHADERMANAGER_H

### 2018-03-16 src/ComputeShaderManager.cpp

@@ -273,3 +273,19 @@ void ComputeShaderManager::clumpCurvesOperator(unsigned int curvesSSBO, unsigned
 	glUniform1ui(glGetUniformLocation(clumpCurvesShader.getID(), "u_mode"), 1);
 	glDispatchCompute((int(curveCount) / 128) + 1, 1, 1);
 }
+
+//----------------------------------------------------------------------------------------------------------------------
+
+void ComputeShaderManager::startTimer()
+{
+	m_timer = std::chrono::system_clock::now();
+}
+
+//----------------------------------------------------------------------------------------------------------------------
+
+void ComputeShaderManager::endTimer()
+{
+	auto endTimer = std::chrono::system_clock::now();
+	auto elapsed = std::chrono::duration_cast<std::chrono::milliseconds>(endTimer - m_timer);
+	std::cout << "timer measured " << elapsed.count() << "ms\n";
+}

### 2018-03-16 src/ObjMeshModel.cpp

@@ -11,6 +11,7 @@
 #include <sstream>
 #include <boost/algorithm/string/predicate.hpp>
 
+#include "ComputeShaderManager.h"
 
 ObjMeshModel::ObjMeshModel() :
 	m_label(new QLabel("Double click to load obj"))
@@ -58,6 +59,8 @@ bool ObjMeshModel::eventFilter(QObject *object, QEvent *event)
 				nullptr, tr("Load Obj"),
 				QDir::currentPath() + "/meshes",
 				tr("Mesh Files (*.obj)"));
+
+			ComputeShaderManager::getInstance()->startTimer();
 
 			m_label->setText(fileName);
 

### 2018-03-16 src/RandomDistributorModel.cpp

@@ -5,6 +5,7 @@
 #include "CurvesData.h"
 
 #include <random>
+// #include <chrono>
 
 RandomDistributorModel::RandomDistributorModel() :
 	m_embedded(new QWidget()),
@@ -44,6 +45,9 @@ void RandomDistributorModel::meshChanged()
 
 void RandomDistributorModel::distribute()
 {
+// 	auto startDistributionTimer = std::chrono::system_clock::now();
+// 	auto endDistributionTimer = startDistributionTimer;
+
 	if (m_ui->modeCheckBox->isChecked()) // GPU mode checked
 	{
 		// this function needs to create & resize m_curvesSSBOID
@@ -56,6 +60,8 @@ void RandomDistributorModel::distribute()
 			m_ui->lengthSpinBox->value(),
 			m_ui->variationSpinBox->value()
 			);
+
+		// endDistributionTimer = std::chrono::system_clock::now();
 	}
 	else
 	{
@@ -89,6 +95,7 @@ void RandomDistributorModel::distribute()
 				{
 					float baryA = distribution(generator);
 					float baryB = distribution(generator);
+					float randLength = m_ui->lengthSpinBox->value() * glm::mix((1.0 - m_ui->variationSpinBox->value()), 1.0, distribution(generator));
 
 					if ((baryA + baryB) > 1.0)
 					{
@@ -109,16 +116,21 @@ void RandomDistributorModel::distribute()
 
 					for (int j = 0; j < 5; ++j)
 					{
-						curve.vertices[j] = randPos + ((float(j) / 5.0f) * randNorm * float(m_ui->lengthSpinBox->value()));
+						curve.vertices[j] = randPos + ((float(j) / 5.0f) * randNorm * randLength);
 					}
-					curve.length = m_ui->lengthSpinBox->value();
+					curve.length = randLength;
 
 					m_curves.m_curves.push_back(curve);
 					break;
 				}
 			}
 		}
+
+		// endDistributionTimer = std::chrono::system_clock::now();
 	}
+
+	// auto elapsed = std::chrono::duration_cast<std::chrono::milliseconds>(endDistributionTimer - startDistributionTimer);
+	// std::cout << "distribution of "<<m_ui->countSpinBox->value()<<" took: " << elapsed.count() << "ms\n";
 }
 
 float RandomDistributorModel::calculateMeshArea() const

### 2018-03-16 src/Viewport.cpp

@@ -6,6 +6,7 @@
 #include <iostream>
 
 #include "RenderableManager.h"
+#include "ComputeShaderManager.h"
 
 //----------------------------------------------------------------------------------------------------------------------
 
@@ -112,6 +113,8 @@ void Viewport::paintGL()
 	RenderableManager::getInstance()->constructRenderables();
 	loadMatricesToRenderManager();
 	RenderableManager::getInstance()->drawRenderables();
+
+	ComputeShaderManager::getInstance()->endTimer();
 }
 
 //----------------------------------------------------------------------------------------------------------------------

### 2018-03-16 docs/paper.tex
more docs & shader tweak
more on docs

@@
 Introduction Text.
+\subsection{Related Work} \label{sec:related}
+nvidia hairworks
 
@@
 
-\begin{enumerate}
-\item Generate a uniform random number in the range [0.0 - 1.0].
-\end{enumerate}
+% \begin{enumerate}
+% \item Generate a uniform random number in the range [0.0 - 1.0].
+% \end{enumerate}
 
+% \begin{algorithm}
+% \begin{algorithmic}[1]
+% \Procedure{Naive Random Distributionr}{}
+% \For{k to k}
+% \EndFor
+% \EndProcedure
+% \end{algorithmic}
+% \end{algorithm}
+
+\subsubsection{Compute Shaders} \label{sec:compute shaders}
+
+In order to get this distribution process running on the GPU, I needed to allocate and store data for both the Mesh and Curves in OpenGL buffers, for later use in a compute shader.
 The most common way of storing OBJ data in OpenGL buffers, as described in \cite{objLoading}, is to use a \verb|GL_ARRAY_BUFFER| object to store the vertex data and a \verb|GL_ELEMENT_ARRAY_BUFFER| to store the indices of the \verb|GL_ARRAY_BUFFER| that make up each face.
@@
 Whilst this is less efficient, I found that the overhead in GPU memory usage was negligible, and that this method allowed for much simpler usage in a compute shader.
+Allocating a buffer for the Curves data was much simpler as it consisted of an array of 'Curve' structs, each consisting of 5 vertices.
+Listing~\ref{lst:buffers} shows how these buffers were accessed from within the compute shaders; note how all of the \texttt{vec3} values have been padded to \texttt{vec4}, as the \texttt{std430} layout qualifier requires an alignment of 16 bytes.
+
+\begin{lstlisting}[language=GLSL, label={lst:buffers}, caption={Extract from the Distributor compute shader, showing how the Mesh and Curves buffers are accessed.}]
+struct Face
+{
+	vec4 position[3];
+	vec4 normal[3];
+	vec4 uv[3];
+};
+
+struct Curve
+{
+	vec4 position[5];
+};
+
+layout (std430, binding = 0) buffer facesBuffer
+{
+	Face faces[];
+};
+
+layout (std430, binding = 1) buffer curvesBuffer
+{
+	Curve curves[];
+};
+\end{lstlisting}
+
+\subsubsection{Curve Operators} \label{sec:operators}
 
@@
 
+\begin{figure}[htbp]\centering
+\begin{center}
+\begin{tabular}{||c||c|c|c||}
+\hline
+Mesh Triangle Count & Primitive Count & CPU Time (ms) & GPU Time (ms)\\
+\hline
+\hline
+16 (Triangle Strip) & 10,000 & 2 - 5 & 1 - 2\\
+\hline
+ - & 100,000 & 18 - 41 & 8 - 24\\
+\hline
+ - & 1,000,000 & 120 - 150 & 77 - 110\\
+\hline
+4968 (Stanford Bunny) & 10,000 & 173 - 211 & 1 - 2\\
+\hline
+ - & 100,000 & 1718 - 1756 & 12 - 30\\
+\hline
+ - & 1,000,000 & 17218 - 17259 & 82 - 112\\
+\hline
+\end{tabular}
+\caption{\label{figure:distributionPerformance} Performance metrics showing time taken to distribute varying numbers of primitives, on meshes of varying triangle counts.
+The system used had an Intel Core i7-6700 CPU, and an NVIDIA GTX1080 GPU.}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]\centering
+\begin{center}
+\begin{tabular}{||c|c||}
+\hline
+Mesh Triangle Count & GPU Time (ms)\\
+\hline
+\hline
+16 (Triangle Strip) & 18 - 20\\
+\hline
+4968 (Stanford Bunny) & 33 - 45\\
+\hline
+14067 (Utah Teapot) & 76 - 92\\
+\hline
+\end{tabular}
+\caption{\label{figure:fullPerformance} Performance metrics showing the time taken to perform a full graph evaluation on a complex setup, as seen in INSERT PIC, for meshes of varying triangle counts.}
+\end{center}
+\end{figure}
+
 \section{Conclusion} \label{sec:conclusion}

### 2018-03-16 docs/paper.tex
more on docs

@@
 
-The most popular of these is XGen, originally developed by Walt Disney Animation Studios, which has been included in Autodesk's Maya since 2014.
+The most popular of these is XGen, originally developed by Walt Disney Animation Studios \cite{xgen}, which has been included in Autodesk's Maya since 2014.
 XGen uses a 'Collection' and 'Description' paradigm; Descriptions store user controlled parameters which affect fur primitives (curves for example) on a select area of a mesh, and Collections act as a container for these Descriptions.
@@
 
-Offerering even even more flexibility, recent versions of SideFX's Houdini include hair and fur operator nodes, which create predefined node networks for the generation of hair and fur.
+Offerering even more flexibility, recent versions of SideFX's Houdini include hair and fur operator nodes, which create predefined node networks for the generation of hair and fur.
 These networks use standard Houdini nodes internally, so experienced Houdini users are intuitively able to create the visual outcome they desire, using tools they already know.
@@
 These parameters could potentially be controlled by texture inputs.
-	\item Operators - These are responsible for manipulation of the curves to achieve the desired look. Example operators could provide bending, clumping, or randomisation of input curves.
+	\item Operators - These are responsible for manipulation of the curves to achieve the desired look.
+Example operators could provide bending, clumping, or randomisation of input curves.
+These should behave much like the 'Modifiers' in XGen.
 	\item Renderers - These are responsible for the rendering of curves into the application viewport.
@@
 
+Once curve primitives have been generated, they need to be passed onto curve operators, which perform most of the work towards creating an appealing visual output.
+I decided to create an abstract curve operator node, much like I had with the distributor node, which contains a virtual function for performing the curve manipulation.
+
 \subsection{User Testing} \label{sec:usertesting}
@@
 \section{Results} \label{sec:results}
+
+\begin{figure}[htbp]\centering
+\includegraphics[width=1.0\linewidth]{images/test_0007}
+\caption{\label{figure:screenshot} An example of the visual output that the system is capable of producing.}
+\end{figure}
+
 Results Text.
@@
 \end{tabular}
-\caption{\label{figure:fullPerformance} Performance metrics showing the time taken to perform a full graph evaluation on a complex setup, as seen in INSERT PIC, for meshes of varying triangle counts.}
+\caption{\label{figure:fullPerformance} Performance metrics showing the time taken to perform a full graph evaluation on a complex setup, as seen in Figure~\ref{figure:screenshot}, for meshes of varying triangle counts.}
 \end{center}

### 2018-03-16 docs/paper.tex
more on docs

@@
 
+layout (std430, binding = 0) buffer facesBuffer
+{
+	Face faces[];
+};
+
 struct Curve
@@
 	vec4 position[5];
-};
-
-layout (std430, binding = 0) buffer facesBuffer
-{
-	Face faces[];
 };
@@
 I decided to create an abstract curve operator node, much like I had with the distributor node, which contains a virtual function for performing the curve manipulation.
+This allows for multiple different curve operators to be integrated into the system with ease, however for the scope of this assignment I chose only to implement curve operators to provide bending, application of noise, and clumping functionality.
 
@@
 \caption{\label{figure:distributionPerformance} Performance metrics showing time taken to distribute varying numbers of primitives, on meshes of varying triangle counts.
-The system used had an Intel Core i7-6700 CPU, and an NVIDIA GTX1080 GPU.}
+An Intel Core i7-6700 CPU and an NVIDIA GTX1080 GPU were used for testing.}
 \end{center}
@@
 
+\section{Future Work} \label{sec:futurework}
+
+Unfortunately my system lacks the ability to export curves data, making it virtually unusable within a production pipeline in its current state.
+
 \bibliographystyle{acmsiggraph}

### 2018-03-17 docs/paper.tex
more on docs

@@
 
+\subsubsection{Rendering} \label{sec:rendering}
+
+By default, 'Mesh' and 'Curves' objects present in the node graph are not rendered in the viewport.
+For example, the user may want to distribute curves from hidden 'scalp' geometry, or use a set of hidden curves as clumping attractors.
+Therefore it was necessary to implement a pair of rendering nodes that would take a 'Mesh' or 'Curves' object as an input, and provide shading controls for rendering them in the viewport.
+
+Whilst rendering and shading triangulated meshes is relatively straightforward, as described in \cite{objLoading}, there are multiple different approaches one could use to render curves.
+The technique I chose to implement relies on drawing an empty VAO multiple times, and using \texttt{gl\_PrimitiveIDIn} from within a geometry shader to determine which curve from the bound \texttt{GL\_SHADER\_STORAGE\_BUFFER} should be drawn.
+Figure~\ref\{} shows how this was achieved.
+
+\begin{lstlisting}[language=GLSL, label={lst:buffers}, caption={Extract from the Distributor compute shader, showing how the Mesh and Curves buffers are accessed.}]
+struct Face
+{
+	vec4 position[3];
+	vec4 normal[3];
+	vec4 uv[3];
+};
+
+layout (std430, binding = 0) buffer facesBuffer
+{
+	Face faces[];
+};
+
+struct Curve
+{
+	vec4 position[5];
+};
+
+layout (std430, binding = 1) buffer curvesBuffer
+{
+	Curve curves[];
+};
+\end{lstlisting}
+
+% I opted to make use of OpenGL's geometry shaders, drawing an empty VAO multiple times, and using \texttt{gl\_PrimitiveIDIn} from within a geometry shader to determine which curve from the bound \texttt{GL\_SHADER\_STORAGE\_BUFFER} should be drawn.
+
 \subsection{User Testing} \label{sec:usertesting}

### 2018-03-18 docs/paper.tex
added appendix

@@
 
-\begin{lstlisting}[language=GLSL, label={lst:buffers}, caption={Extract from the Distributor compute shader, showing how the Mesh and Curves buffers are accessed.}]
-struct Face
-{
-	vec4 position[3];
-	vec4 normal[3];
-	vec4 uv[3];
-};
-
-layout (std430, binding = 0) buffer facesBuffer
-{
-	Face faces[];
-};
-
-struct Curve
-{
-	vec4 position[5];
-};
-
-layout (std430, binding = 1) buffer curvesBuffer
-{
-	Curve curves[];
-};
-\end{lstlisting}
-
 \subsubsection{Curve Operators} \label{sec:operators}
@@
 The technique I chose to implement relies on drawing an empty VAO multiple times, and using \texttt{gl\_PrimitiveIDIn} from within a geometry shader to determine which curve from the bound \texttt{GL\_SHADER\_STORAGE\_BUFFER} should be drawn.
-Figure~\ref\{} shows how this was achieved.
-
-\begin{lstlisting}[language=GLSL, label={lst:buffers}, caption={Extract from the Distributor compute shader, showing how the Mesh and Curves buffers are accessed.}]
-struct Face
-{
-	vec4 position[3];
-	vec4 normal[3];
-	vec4 uv[3];
-};
-
-layout (std430, binding = 0) buffer facesBuffer
-{
-	Face faces[];
-};
-
-struct Curve
-{
-	vec4 position[5];
-};
-
-layout (std430, binding = 1) buffer curvesBuffer
-{
-	Curve curves[];
-};
-\end{lstlisting}
+Listings~\ref{lst:rendercpp} and ~\ref{lst:renderglsl} shows how this was achieved.
 
@@
 
+\newpage
+\appendix
+
+\begin{lstlisting}[language=GLSL, label={lst:buffers}, caption={Extract from the Distributor compute shader, showing how the Mesh and Curves buffers are accessed.}]
+struct Face
+{
+	vec4 position[3];
+	vec4 normal[3];
+	vec4 uv[3];
+};
+
+layout (std430, binding = 0) buffer facesBuffer
+{
+	Face faces[];
+};
+
+struct Curve
+{
+	vec4 position[5];
+};
+
+layout (std430, binding = 1) buffer curvesBuffer
+{
+	Curve curves[];
+};
+\end{lstlisting}
+
+\begin{lstlisting}[language=C++, label={lst:rendercpp}, caption={Extract from the C++ application code showing how Curves are rendered procedurally.
+\texttt{emptyVAO} is an empty Vertex Array Object, \texttt{curvesSSBO} is a \texttt{GL\_SHADER\_STORAGE\_BUFFER} containing vertex data for multiple curves, and \texttt{indices} is the number of curves in the buffer.}]
+glBindVertexArray(emptyVAO);
+glBindBufferBase(GL_SHADER_STORAGE_BUFFER, 0, curvesSSBO);
+glDrawArrays(GL_POINTS, 0, indices);
+\end{lstlisting}
+
+\begin{lstlisting}[language=GLSL, label={lst:renderglsl}, caption={Extract from the curves rendering geometry shader, showing how vertex positions were processed from data stored in a bound \texttt{GL\_SHADER\_STORAGE\_BUFFER}.}]
+#version 430 core
+layout(points) in;
+layout(line_strip, max_vertices = 5) out;
+
+struct Curve
+{
+	vec4 position[5];
+};
+
+layout (std430, binding = 0) buffer curvesBuffer
+{
+	Curve curves[];
+};
+
+uniform mat4 MVP;
+
+void main()
+{
+	for (int i = 0; i < 5; ++i)
+	{
+		gl_Position = MVP * vec4(curves[gl_PrimitiveIDIn].position[i].xyz, 1.0);
+		EmitVertex();
+	}
+	EndPrimitive();
+}
+\end{lstlisting}
+
 \end{document}

### 2018-03-18 docs/paper.tex
more on docs

@@
 Introduction Text.
-\subsection{Related Work} \label{sec:related}
+\subsection{Related work} \label{sec:related}
 nvidia hairworks
 
-\subsection{Existing Solutions} \label{sec:existing}
-Whilst many larger companies opt to develop their own software for handling fur, such as MPC's Furtility \cite{furtility}, there are a number of commercially available solutions that are also widely used in production.
+\subsection{Existing solutions} \label{sec:existing}
+Whilst many larger companies opt to develop their own software for handling fur, such as MPC's Furtility~\cite{furtility}, there are a number of commercially available solutions that are also widely used in production.
 
-The most popular of these is XGen, originally developed by Walt Disney Animation Studios \cite{xgen}, which has been included in Autodesk's Maya since 2014.
+The most popular of these is XGen, originally developed by Walt Disney Animation Studios~\cite{xgen}, which has been included in Autodesk's Maya since 2014.
 XGen uses a 'Collection' and 'Description' paradigm; Descriptions store user controlled parameters which affect fur primitives (curves for example) on a select area of a mesh, and Collections act as a container for these Descriptions.
@@
 
-Another commercially available solution is Peregrine*Labs' Yeti \cite{yeti}, which is also offered as a plugin within Autodesk's Maya.
+Another commercially available solution is Peregrine*Labs' Yeti~\cite{yeti}, which is also offered as a plugin within Autodesk's Maya.
 Yeti offers a more procedural approach, giving the user a node graph interface in which to construct the desired look from individual components.
@@
 
-\subsection{GPU Acceleration} \label{sec:existingGPU}
+\subsection{GPU acceleration} \label{sec:existingGPU}
 Whilst it is not clear to what extent each of these solutions make use of GPU acceleration, they all perform fast enough to allow the user to manipulate them interactively.
@@
 I wanted to include a node-graph style interface within my application, as it is commonly used within existing artist tools (Autodesk Maya, Unreal Engine), and would encourage modularity within my API.
-Qt does not natively provide this kind of interface, so I made use of NodeEditor \cite{Pinaev2017}, an existing Qt-based library that provides this functionality.
+Qt does not natively provide this kind of interface, so I made use of NodeEditor~\cite{Pinaev2017}, an existing Qt-based library that provides this functionality.
 
@@
 
-\subsubsection{Mesh Loading} \label{sec:objLoading}
+\subsubsection{Mesh loading} \label{sec:objLoading}
 % The first problem I tackled was that of loading mesh geometry into the system, and storing it in a Shader Storage Buffer Object (SSBO) for later use in a compute shader.
@@
 
-\subsubsection{Curve Distributors} \label{sec:distributors}
+\subsubsection{Curve distributors} \label{sec:distributors}
 
@@
 Whilst these kinds of methods produce good results, they can be inefficient as in certain situations many samples may be generated and rejected, before a sample is accepted.
-An optimisation for this is described in \textit{Dart Throwing on Surfaces} \cite{dartThrowing}, which relies on removing areas of the mesh that are already covered by a sample, by means of triangle subdivision.
+An optimisation for this is described in \textit{Dart Throwing on Surfaces}~\cite{dartThrowing}, which relies on removing areas of the mesh that are already covered by a sample, by means of triangle subdivision.
 
 Unfortunately, parallelizing dart thowing methods is not trivial, making them unsuitable for my desired usage of OpenGL compute shaders.
-Whilst sampling methods suitable for parallel processing do exist, such as those described in \textit{Parallel Poisson Disk Sampling with Spectrum Analysis on Surfaces} \cite{parallelPoisson}, I instead opted to implement a naive random samling method as it was much easier to parallelize.
+Whilst sampling methods suitable for parallel processing do exist, such as those described in \textit{Parallel Poisson Disk Sampling with Spectrum Analysis on Surfaces}~\cite{parallelPoisson}, I instead opted to implement a naive random sampling method as it was much easier to parallelize.
 
@@
 
-\subsubsection{Compute Shaders} \label{sec:compute shaders}
+\subsubsection{Compute shaders} \label{sec:compute shaders}
 
 In order to get this distribution process running on the GPU, I needed to allocate and store data for both the Mesh and Curves in OpenGL buffers, for later use in a compute shader.
-The most common way of storing OBJ data in OpenGL buffers, as described in \cite{objLoading}, is to use a \verb|GL_ARRAY_BUFFER| object to store the vertex data and a \verb|GL_ELEMENT_ARRAY_BUFFER| to store the indices of the \verb|GL_ARRAY_BUFFER| that make up each face.
+The most common way of storing OBJ data in OpenGL buffers, as described in~\cite{objLoading}, is to use a \verb|GL_ARRAY_BUFFER| object to store the vertex data and a \verb|GL_ELEMENT_ARRAY_BUFFER| to store the indices of the \verb|GL_ARRAY_BUFFER| that make up each face.
 However for ease of use in compute shaders I decided instead to store each face in a \verb|GL_SHADER_STORAGE_BUFFER|, where each face consists of three sets of positions, normals, and texture coordinates.
@@
 
-\subsubsection{Curve Operators} \label{sec:operators}
+\subsubsection{Curve operators} \label{sec:operators}
 
@@
 
-Whilst rendering and shading triangulated meshes is relatively straightforward, as described in \cite{objLoading}, there are multiple different approaches one could use to render curves.
+Whilst rendering and shading triangulated meshes is relatively straightforward, as described in~\cite{objLoading}, there are multiple different approaches one could use to render curves.
 The technique I chose to implement relies on drawing an empty VAO multiple times, and using \texttt{gl\_PrimitiveIDIn} from within a geometry shader to determine which curve from the bound \texttt{GL\_SHADER\_STORAGE\_BUFFER} should be drawn.
 Listings~\ref{lst:rendercpp} and ~\ref{lst:renderglsl} shows how this was achieved.
+The use geometry shaders also allowed me to alternate between drawing the curves using line primitives, or to draw the curves as forward facing triangle strips.
+% which I refer to as 'ribbons'.
+
+For shading of the meshes I opted to implement a physically based shading model, as decribed in~\cite{learnPBR}, however some alterations were necessary as my method of rendering curves does not provide usable surface normal information.
+Instead, the normalized direction in which the curve is leading from one vertex to the next, was substituted as the surface normal.
+Then, all of the dot product terms in the shading calculations which use the surface normal, were inversed.
+The result roughly approximates the shading qualities of thin cylinders by assuming that all vectors that are perpendicular to the direction of the curve are valid surface normals.
+Whilst this technique provides visually pleasing results, it is inaccurate and does not account for properties such as self-shadowing, or light scattering.
 
@@
 
-\subsection{User Testing} \label{sec:usertesting}
+\subsection{User testing} \label{sec:usertesting}
 Near the end of the development I was able to convince friend and fellow student Alin Bolcas to test the system and provide feedback from an artist's point of view; Alin is an experienced modelling and look development artist who uses a variety of different software packages to create his work, so getting his feedback on the usability of my tool proved extremely useful.
@@
 
+After I had implemented some of the features Alin requested, I decided that the system was in a satisfactory state for the purposes of this assignment.
+The following figure~\ref{figure:screenshots} shows some memes
+
+\begin{figure}[htbp]
+\centering
+\subfloat[]{\includegraphics[width=0.45\linewidth]{images/rick}}
+\hfill
+\subfloat[]{\includegraphics[width=0.45\linewidth]{images/bunny}}
+\caption{\label{figure:screenshots} literal memes}
+\end{figure}
+
 \begin{figure}[htbp]\centering
 \includegraphics[width=1.0\linewidth]{images/test_0007}
-\caption{\label{figure:screenshot} An example of the visual output that the system is capable of producing.}
+\caption{\label{figure:screenshot1} An example of the visual output that the system is capable of producing.}
 \end{figure}
@@
 
-\section{Future Work} \label{sec:futurework}
+\subsection{Future Work} \label{sec:futurework}
 

### 2018-03-18 docs/paper.tex
more on docs

@@
 
-The most popular of these is XGen, originally developed by Walt Disney Animation Studios~\cite{xgen}, which has been included in Autodesk's Maya since 2014.
+The most popular of these is XGen~\cite{xgen}, originally developed by Walt Disney Animation Studios, which has been included in Autodesk's Maya since 2014.
 XGen uses a 'Collection' and 'Description' paradigm; Descriptions store user controlled parameters which affect fur primitives (curves for example) on a select area of a mesh, and Collections act as a container for these Descriptions.
@@
 
-Another commercially available solution is Peregrine*Labs' Yeti~\cite{yeti}, which is also offered as a plugin within Autodesk's Maya.
+Another commercially available solution is Yeti~\cite{yeti}, which is also offered as a plugin within Autodesk's Maya.
 Yeti offers a more procedural approach, giving the user a node graph interface in which to construct the desired look from individual components.
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.49\linewidth]{images/triangleStripDistribution}
+\hfill
+\includegraphics[width=0.45\linewidth]{images/triangleMesh}
+\caption{\label{figure:distribution} Example distribution generated by my naive random distributor, note how the distribution stays relatively uniform despite variance in mesh density.}
+\end{figure}
+
 % \begin{enumerate}
@@
 I decided to create an abstract curve operator node, much like I had with the distributor node, which contains a virtual function for performing the curve manipulation.
-This allows for multiple different curve operators to be integrated into the system with ease, however for the scope of this assignment I chose only to implement curve operators to provide bending, application of noise, and clumping functionality.
+This allows for multiple different curve operators to be integrated into the system with ease, however for the scope of this assignment I chose only to implement curve operators to provide bending and clumping functionality.
 
@@
 For shading of the meshes I opted to implement a physically based shading model, as decribed in~\cite{learnPBR}, however some alterations were necessary as my method of rendering curves does not provide usable surface normal information.
-Instead, the normalized direction in which the curve is leading from one vertex to the next, was substituted as the surface normal.
+Instead the normalized direction in which the curve is leading, from one vertex to the next, was substituted as the surface normal.
 Then, all of the dot product terms in the shading calculations which use the surface normal, were inversed.
@@
 After I had implemented some of the features Alin requested, I decided that the system was in a satisfactory state for the purposes of this assignment.
-The following figure~\ref{figure:screenshots} shows some memes
+The following Figures~\ref{figure:screenshots} and ~\ref{figure:screenshot_gui} show examples of the kind of visual output my system is capable of producing, and the accompanying node graph setup used to create them.
 
@@
 \centering
-\subfloat[]{\includegraphics[width=0.45\linewidth]{images/rick}}
+\includegraphics[width=0.45\linewidth]{images/rick}
 \hfill
-\subfloat[]{\includegraphics[width=0.45\linewidth]{images/bunny}}
-\caption{\label{figure:screenshots} literal memes}
+\includegraphics[width=0.45\linewidth]{images/bunny}
+\caption{\label{figure:screenshots} Examples of visual output that my system is capable of producing.}
 \end{figure}
@@
 \begin{figure}[htbp]\centering
-\includegraphics[width=1.0\linewidth]{images/test_0007}
-\caption{\label{figure:screenshot1} An example of the visual output that the system is capable of producing.}
+\includegraphics[width=1.0\linewidth]{images/nodes}
+\caption{\label{figure:screenshot_gui} The accompanying node graph setup used to create the images in Figure~\ref{figure:screenshots}.}
 \end{figure}
-
-Results Text.
 
@@
 
-insert stats comparing CPU/GPU compute
+Figures~\ref{figure:distributionPerformance} and~\ref{figure:fullPerformance} show that my system is performant enough to provide real-time feedback, even when dense meshes and high primitive counts are used.
+Whilst Figure~\ref{figure:distributionPerformance} shows performance gains by utilisizing the GPU over the CPU, note that my CPU implementation is single-threaded, and further performance gains could be made by utilizing multithreading.
+
+It is worth noting that my system is quite inefficient in regards to GPU memory consumption; The node graph setup in Figure~\ref{figure:screenshot_gui} uses 322 MiB of VRAM.
+This is a result of storing the state of each \texttt{GL\_SHADER\_STORAGE\_BUFFER} in GPU memory, for each of the curve operators used.
+Whilst this avoids the need for a full graph evaluation each time a parameter is changed, it effectively doubles VRAM usage each time a curve operator is added.
 
@@
 
+% \begin{figure}[htbp]\centering
+% \begin{center}
+% \begin{tabular}{||c|c||}
+% \hline
+% Mesh Triangle Count & GPU Time (ms)\\
+% \hline
+% \hline
+% 16 (Triangle Strip) & 18 - 20\\
+% \hline
+% 4968 (Stanford Bunny) & 33 - 45\\
+% \hline
+% 14067 (Utah Teapot) & 76 - 92\\
+% \hline
+% \end{tabular}
+% \caption{\label{figure:fullPerformance} Performance metrics showing the time taken to perform a full graph evaluation on a complex setup, as seen in Figure~\ref{figure:screenshot_gui}, for meshes of varying triangle counts.}
+% \end{center}
+% \end{figure}
+
 \begin{figure}[htbp]\centering
@@
 \hline
-Mesh Triangle Count & GPU Time (ms)\\
+Mesh Triangle Count & Evaluation Time (ms)\\
 \hline
 \hline
-16 (Triangle Strip) & 18 - 20\\
+16 (Triangle Strip) & 69 - 84\\
 \hline
-4968 (Stanford Bunny) & 33 - 45\\
+4968 (Stanford Bunny) & 81 - 112\\
 \hline
-14067 (Utah Teapot) & 76 - 92\\
+14067 (Utah Teapot) & 138 - 228\\
 \hline
 \end{tabular}
-\caption{\label{figure:fullPerformance} Performance metrics showing the time taken to perform a full graph evaluation on a complex setup, as seen in Figure~\ref{figure:screenshot}, for meshes of varying triangle counts.}
+\caption{\label{figure:fullPerformance} Performance metrics showing the time taken to perform a full graph evaluation on a complex setup, as seen in Figure~\ref{figure:screenshot_gui}, for meshes of varying triangle counts.
+Note that over 500,000 primitives are being processed in this setup.}
 \end{center}
@@
 
+Overall I am pleased with the visual quality and interactivity of my system, but I believe further work is necessary in order to maximise it's performance and usability.
+
+% Currently my system performs all of its computation using OpenGL compute shaders, though better results may be attainable using other APIs such as CUDA or OpenCL.
+
 \subsection{Future Work} \label{sec:futurework}
@@
 Unfortunately my system lacks the ability to export curves data, making it virtually unusable within a production pipeline in its current state.
+A viable file format for this could be \textit{Alembic}~\cite{alembic}, as it natively supports the storage of parametric curves.
+Integrating alembic support into my system could also allow for animated mesh geometry to be used.
+
+Another desirable addition to my system would be viewport manipulators, similar to the 'Groomable Splines' feature in XGen.
+This could allow me to implement techniques such as those described in Skippy~\cite{skippy}, which would provide my system with a unique user experience compared to commercially available systems.
+
+Whilst not necessary for usage in a broader pipeline, implementing a physically plausible hair shading model such as those described in \cite{Marschner}, would greatly improve the visual accuracy of my viewport renderings.
+
+\newpage
 
@@
 	vec4 normal[3];
-	vec4 uv[3];
+	vec4 uv[3];{}
 };

### 2018-03-18 docs/paper.tex
more on docs

@@
 \subsection{Related work} \label{sec:related}
+
 nvidia hairworks
@@
 
-Overall I am pleased with the visual quality and interactivity of my system, but I believe further work is necessary in order to maximise it's performance and usability.
+Overall I am pleased with the visual quality and interactivity of my system; currently it provides the basic functionality required to visualise simple hair and fur content, whether it be for a VFX or real-time usage scenario.
+However, I believe further work is necessary in order to maximise it's performance and usability within a broader production pipeline.
 

### 2018-03-18 docs/paper.tex
more on docs

@@
 \section{Introduction} \label{sec:introduction}
-Introduction Text.
+Generating and rendering hair and fur content for use in visual effects and games is typically a very computationally heavy task, due to the inherent complexity of handling thousands if not millions of geometric strands.
+Fortunately, modern hardware provides flexibility in regards to accelerating such tasks, which we are able to leverage though the use of hardware acceleration APIs such as OpenCL and CUDA.
+
 \subsection{Related work} \label{sec:related}
-
-nvidia hairworks
-
-\subsection{Existing solutions} \label{sec:existing}
 Whilst many larger companies opt to develop their own software for handling fur, such as MPC's Furtility~\cite{furtility}, there are a number of commercially available solutions that are also widely used in production.
@@
 Whilst it is not clear to what extent each of these solutions make use of GPU acceleration, they all perform fast enough to allow the user to manipulate them interactively.
-I therefore decided to monitor the CPU and GPU usage using \verb|htop| and \verb|nvidia-smi| respectively, before and after adding XGen content to a simple maya scene, to determine where the majority of data was being computed.
+However, software such as the HairWorks \cite{hairworks} API and TressFX \cite{tressfx} library are developed primarily for use in games, so they are designed to utilise the GPU as much as possible.
+These solutions are able to perform both simulation and rendering of hair in real-time, providing interactivity that exceeds that of comparable software designed for VFX usage.
+It is worth noting that these tools do not provide a method for authoring hair and fur content, but instead designed to use load content generated by 3rd party authoring tools.
 
-After applying the default XGen description onto a simple sphere with 760 triangles, there was no noticeable increase in both GPU utilisation and GPU memory usage.
+\begin{figure}[htbp]\centering
+\includegraphics[width=1.0\linewidth]{images/hairworks}
+\caption{\label{figure:hairworks} NVIDIA's HairWorks API is used in a handful of modern games, such as \textit{The Witcher 3} \protect\cite{witcher}, to provide realistic and interactive hair rendering.}
+\end{figure}
+
+% I therefore decided to monitor the CPU and GPU usage using \verb|htop| and \verb|nvidia-smi| respectively, before and after adding XGen content to a simple maya scene, to determine where the majority of data was being computed.
+
+% After applying the default XGen description onto a simple sphere with 760 triangles, there was no noticeable increase in both GPU utilisation and GPU memory usage.
 Increasing the density of the primitives from 1.0 to 10000.0 saw an increase of VRAM usage of roughly 350MiB, which I would attribute to storing the primitives in a buffer on the GPU, but there was no discernable increase in GPU utilisation.
@@
 
-However, when applying modifiers to the XGen description I noticed a small increase of roughly 2MiB in VRAM usage for the first modifier, yet no further increases for modifiers added after that.
+% However, when applying modifiers to the XGen description I noticed a small increase of roughly 2MiB in VRAM usage for the first modifier, yet no further increases for modifiers added after that.
 As this increase is quite small I would attribute it to storing the modifier attributes on the GPU, though I still noticed spikes in CPU usage when recalculating the XGen description once the modifier was added, which makes it inconclusive as to whether XGen uses any kind of GPU acceleration.
@@
 \hline
-Mesh Triangle Count & Primitive Count & CPU Time (ms) & GPU Time (ms)\\
+Mesh Triangle Count & Primitive Count & CPU (ms) & GPU (ms)\\
 \hline
@@
 \end{tabular}
-\caption{\label{figure:distributionPerformance} Performance metrics showing time taken to distribute varying numbers of primitives, on meshes of varying triangle counts.
+\caption{\label{figure:distributionPerformance} Performance metrics comparing time taken to distribute varying numbers of primitives, on meshes of varying densities, using the CPU and GPU.
 An Intel Core i7-6700 CPU and an NVIDIA GTX1080 GPU were used for testing.}
@@
 
-Whilst not necessary for usage in a broader pipeline, implementing a physically plausible hair shading model such as those described in \cite{Marschner}, would greatly improve the visual accuracy of my viewport renderings.
+Whilst not necessary for usage in a offline pipeline, implementing a physically plausible hair shading model based on the techniques described in \cite{Marschner}, would greatly improve the visual accuracy of my viewport renderings.
 

### 2018-03-19 docs/paper.tex
spell checked, need to proof read

@@
 XGen uses a 'Collection' and 'Description' paradigm; Descriptions store user controlled parameters which affect fur primitives (curves for example) on a select area of a mesh, and Collections act as a container for these Descriptions.
-There are two primary methods offered for distributing primitives on a mesh, Guide Curves can be used to manually control the placement and shape of primitives by placing 'Guides' on the mesh surface, whereas 'Groomable Splines' randomly distirbutes splines over the mesh surface which can be manipulated in real time using viewport manipulators that mimic the brushing and styling of fur.
+There are two primary methods offered for distributing primitives on a mesh, Guide Curves can be used to manually control the placement and shape of primitives by placing 'Guides' on the mesh surface, whereas 'Groomable Splines' randomly distributes splines over the mesh surface which can be manipulated in real time using viewport manipulators that mimic the brushing and styling of fur.
 In both modes, 'modifiers' can be added to the modifier stack to further adjust the visual outcome, adding features such as clumping, noise, and coiling.
@@
 
-Offerering even more flexibility, recent versions of SideFX's Houdini include hair and fur operator nodes, which create predefined node networks for the generation of hair and fur.
+Offering even more flexibility, recent versions of SideFX's Houdini include hair and fur operator nodes, which create predefined node networks for the generation of hair and fur.
 These networks use standard Houdini nodes internally, so experienced Houdini users are intuitively able to create the visual outcome they desire, using tools they already know.
@@
 
-% I therefore decided to monitor the CPU and GPU usage using \verb|htop| and \verb|nvidia-smi| respectively, before and after adding XGen content to a simple maya scene, to determine where the majority of data was being computed.
-
-% After applying the default XGen description onto a simple sphere with 760 triangles, there was no noticeable increase in both GPU utilisation and GPU memory usage.
-Increasing the density of the primitives from 1.0 to 10000.0 saw an increase of VRAM usage of roughly 350MiB, which I would attribute to storing the primitives in a buffer on the GPU, but there was no discernable increase in GPU utilisation.
-Increasing the density further to 10,000 saw a noticable spike in CPU usage, hitting 100 percent across two threads, leading me to believe that the primitive distribution computation is being done on the CPU.
-
-% However, when applying modifiers to the XGen description I noticed a small increase of roughly 2MiB in VRAM usage for the first modifier, yet no further increases for modifiers added after that.
-As this increase is quite small I would attribute it to storing the modifier attributes on the GPU, though I still noticed spikes in CPU usage when recalculating the XGen description once the modifier was added, which makes it inconclusive as to whether XGen uses any kind of GPU acceleration.
-
-% on maya load: CPU up to 8 percent when moving mouse GPU very low but 40 percent when moving viewport 42MiB
-% before xgen on sphere 36mib
-% xgen on sphere density 1.0 36mib GPU
-% xgen on sphere density 100.0 40mib GPU
-% xgen CPU spiked up to 100 on a single thread when changing density
-% xgen on sphere density 1000.0 56mib GPU
-% xgen noise on sphere density 1000.0 63mib GPU, CPU spiked but only to 20%
-% xgen on sphere density to 10,000 caused spike on 2 threads
-
 \section{Implementation} \label{sec:implementation}
-From my research it was clear that there is an apparent lack of standalone software for authoring fur; it makes sense that existing fur software are developed as software plugins, allowing the tool to make use of their host DCCs existing functionality.
+From my research it was clear that there is an apparent lack of standalone software for authoring fur; it makes sense that existing fur software are developed as software plugins, as this allows the tool to make use of the host digital content creation (DCC) package's existing functionality.
 
 However, by developing the system as a simple API one could make it DCC agnostic, which is why I chose to develop my system as a pairing of a simple API with a standalone tool which interfaces with it.
-By developing a simple API it could allow for future implementations as plugins within DCCs, providing the data exchange formats are compatible with the host DCC.
+By developing a simple API it could allow for a future implementation as a plugin within a DCC, providing the data exchange formats are compatible with the host.
 
 \subsection{Resources} \label{sec:resources}
-I decided to develop the API side of the fur system using C++, primarly as it most commonly used when developing computationally heavy artist tools, but also because it is the language I am most comfortable using.
-I opted to use OpenGL over other APIs (CUDA, OpenCL) for offloading of computation onto the GPU, simply because my final artefact is a tool with a graphical interface, and OpenGL is capable of handling both arbitary computation using compute shaders and rendering of geometry simultaneously.
+I decided to develop the API side of the fur system using C++, primarily as it most commonly used when developing computationally heavy artist tools, but also because it is the language I am most comfortable using.
+I opted to use OpenGL over other APIs (CUDA, OpenCL) for offloading of computation onto the GPU, simply because my final artefact is a tool with a graphical interface, and OpenGL is capable of handling both arbitrary computation using compute shaders and rendering of geometry simultaneously.
 
 To handle the user interface I used the Qt framework within C++.
-Qt is commonly used for artist tools within visual effects as it is cross-platform, and applications can be configured to run within other applications that make use of it, such as Autodesk Maya.
+Qt is commonly used for artist tools within visual effects as it is cross-platform, and applications can be configured to run within other applications that make use of it, such as Autodesk's Maya.
 
-I wanted to include a node-graph style interface within my application, as it is commonly used within existing artist tools (Autodesk Maya, Unreal Engine), and would encourage modularity within my API.
+I wanted to include a node-graph style interface within my application, as these are commonly used within existing artist tools (Autodesk Maya, Unreal Engine), and would encourage modularity within my API.
 Qt does not natively provide this kind of interface, so I made use of NodeEditor~\cite{Pinaev2017}, an existing Qt-based library that provides this functionality.
@@
 	\item Renderers - These are responsible for the rendering of curves into the application viewport.
-Example renderers could provide mesh, cuves as lines, or curves as ribbons rendering functionality.
+Example renderers could provide mesh, curves as lines, or curves as ribbons rendering functionality.
 User controllable parameters could provide controls for the shading model in use, as well as control of the base and tip widths when rendering curve ribbons for example.
@@
 I chose to omit this functionality to limit the scope of this assignment, but the ability to export hair curves as alembic would be desirable, as alembic is a widely adopted format that allows for the storage of multiple different types of curve geometry.
-Another candidate file type would be RenderMan's RIB format, which would allow for direct rendering of the hair curves in any RenderMan compliant renderer.
 
@@
 
-In this section I will explain the development of each of the major components, not necessarily in chronological order, but in the order regarding to how their data is passed through the system.
-
-% write about distributing https://stackoverflow.com/questions/9294316/
-% distribute-points-on-mesh-according-to-density
-
-% write about GPU
+In this section I will explain the development of each of the major components, not necessarily in chronological order, but in order regarding to how their data is passed through the system.
 
 \subsubsection{Mesh loading} \label{sec:objLoading}
-% The first problem I tackled was that of loading mesh geometry into the system, and storing it in a Shader Storage Buffer Object (SSBO) for later use in a compute shader.
-I opted to implement a Wavefront OBJ loader as they are (deceptively) simple to parse, consisting of a list of vertex positions, vertex normals, vertex texture coordinates, and faces defined as a list of indices.
-I created an OBJ loader node using the NodeEditor library, which parses
 
@@
 This makes the system somewhat extensible, allowing for different distribution patterns to be implemented with ease.
-The node's data type was also extented to be able to store either 'Curves' objects from the CPU, or an ID for an OpenGL buffer object, with an enumerated value to identify which of these is currently in use.
-This not only allows the distributor to alternate between CPU and GPU computation arbitrarily, but also allows nodes recieving this data to process it accordingly.
+The node's data type was also extended to be able to store either 'Curves' objects from the CPU, or an ID for an OpenGL buffer object, with an enumerated value to identify which of these is currently in use.
+This not only allows the distributor to alternate between CPU and GPU computation arbitrarily, but also allows nodes receiving this data to process it accordingly.
 
-There are many different methods one could use to distribute points on an abritrary mesh, however for replicating creature fur it is desirable to choose a method that distributes primitives at random positions across the surface, yet roughly evenly spaced from neighboring primitives.
-A popular method used to achieve this kind of distribution is referred to as 'Dart Throwing', which relies on checking the distance to neighboring samples after a new sample has been generated, and discarding the sample if the samples are too close.
+There are many different methods one could use to distribute points on an arbitrary mesh, however for replicating creature fur it is desirable to choose a method that distributes primitives at random positions across the surface, yet roughly evenly spaced from neighbouring primitives.
+A popular method used to achieve this kind of distribution is referred to as 'Dart Throwing', which relies on checking the distance to neighbouring samples after a new sample has been generated, and discarding the sample if the samples are too close.
 Whilst these kinds of methods produce good results, they can be inefficient as in certain situations many samples may be generated and rejected, before a sample is accepted.
@@
 
-Unfortunately, parallelizing dart thowing methods is not trivial, making them unsuitable for my desired usage of OpenGL compute shaders.
+Unfortunately, parallelising dart throwing methods is not trivial, making them unsuitable for my desired usage of OpenGL compute shaders.
 Whilst sampling methods suitable for parallel processing do exist, such as those described in \textit{Parallel Poisson Disk Sampling with Spectrum Analysis on Surfaces}~\cite{parallelPoisson}, I instead opted to implement a naive random sampling method as it was much easier to parallelize.
@@
 \end{figure}
-
-% \begin{enumerate}
-% \item Generate a uniform random number in the range [0.0 - 1.0].
-% \end{enumerate}
-
-% \begin{algorithm}
-% \begin{algorithmic}[1]
-% \Procedure{Naive Random Distributionr}{}
-% \For{k to k}
-% \EndFor
-% \EndProcedure
-% \end{algorithmic}
-% \end{algorithm}
 
@@
 
-For shading of the meshes I opted to implement a physically based shading model, as decribed in~\cite{learnPBR}, however some alterations were necessary as my method of rendering curves does not provide usable surface normal information.
+For shading of the meshes I opted to implement a physically based shading model, as described in~\cite{learnPBR}, however some alterations were necessary as my method of rendering curves does not provide usable surface normal information.
 Instead the normalized direction in which the curve is leading, from one vertex to the next, was substituted as the surface normal.
@@
 Whilst this technique provides visually pleasing results, it is inaccurate and does not account for properties such as self-shadowing, or light scattering.
-
-% I opted to make use of OpenGL's geometry shaders, drawing an empty VAO multiple times, and using \texttt{gl\_PrimitiveIDIn} from within a geometry shader to determine which curve from the bound \texttt{GL\_SHADER\_STORAGE\_BUFFER} should be drawn.
 
@@
 Figures~\ref{figure:distributionPerformance} and~\ref{figure:fullPerformance} show that my system is performant enough to provide real-time feedback, even when dense meshes and high primitive counts are used.
-Whilst Figure~\ref{figure:distributionPerformance} shows performance gains by utilisizing the GPU over the CPU, note that my CPU implementation is single-threaded, and further performance gains could be made by utilizing multithreading.
+Whilst Figure~\ref{figure:distributionPerformance} shows performance gains by utilising the GPU over the CPU, note that my CPU implementation is single-threaded, and further performance gains could be made by utilising multithreading.
 
@@
 
-% \begin{figure}[htbp]\centering
-% \begin{center}
-% \begin{tabular}{||c|c||}
-% \hline
-% Mesh Triangle Count & GPU Time (ms)\\
-% \hline
-% \hline
-% 16 (Triangle Strip) & 18 - 20\\
-% \hline
-% 4968 (Stanford Bunny) & 33 - 45\\
-% \hline
-% 14067 (Utah Teapot) & 76 - 92\\
-% \hline
-% \end{tabular}
-% \caption{\label{figure:fullPerformance} Performance metrics showing the time taken to perform a full graph evaluation on a complex setup, as seen in Figure~\ref{figure:screenshot_gui}, for meshes of varying triangle counts.}
-% \end{center}
-% \end{figure}
-
 \begin{figure}[htbp]\centering
@@
 
-% Currently my system performs all of its computation using OpenGL compute shaders, though better results may be attainable using other APIs such as CUDA or OpenCL.
-
 \subsection{Future Work} \label{sec:futurework}
@@
 Integrating alembic support into my system could also allow for animated mesh geometry to be used.
+Another candidate file format would be RenderMan's RIB format, which would allow for direct rendering of the hair curves in any RenderMan compliant renderer.
 

### 2018-03-19 docs/paper.tex
docs tweaks

@@
 \subsection{Related work} \label{sec:related}
-Whilst many larger companies opt to develop their own software for handling fur, such as MPC's Furtility~\cite{furtility}, there are a number of commercially available solutions that are also widely used in production.
+Whilst many larger companies opt to develop their own software for handling fur, such as MPC's \textit{Furtility}~\cite{furtility}, there are a number of commercially available solutions that are also widely used in production.
 
-The most popular of these is XGen~\cite{xgen}, originally developed by Walt Disney Animation Studios, which has been included in Autodesk's Maya since 2014.
+The most popular of these is \textit{XGen}~\cite{xgen}, originally developed by Walt Disney Animation Studios, which has been included in Autodesk's Maya since 2014.
 XGen uses a 'Collection' and 'Description' paradigm; Descriptions store user controlled parameters which affect fur primitives (curves for example) on a select area of a mesh, and Collections act as a container for these Descriptions.
@@
 
-Another commercially available solution is Yeti~\cite{yeti}, which is also offered as a plugin within Autodesk's Maya.
+Another commercially available solution is \textit{Yeti}~\cite{yeti}, which is also offered as a plugin within Autodesk's Maya.
 Yeti offers a more procedural approach, giving the user a node graph interface in which to construct the desired look from individual components.
@@
 Whilst it is not clear to what extent each of these solutions make use of GPU acceleration, they all perform fast enough to allow the user to manipulate them interactively.
-However, software such as the HairWorks \cite{hairworks} API and TressFX \cite{tressfx} library are developed primarily for use in games, so they are designed to utilise the GPU as much as possible.
+However, software such as the \textit{HairWorks}~\cite{hairworks} API and \textit{TressFX}~\cite{tressfx} library are developed primarily for use in games, so they are designed to utilise the GPU as much as possible.
 These solutions are able to perform both simulation and rendering of hair in real-time, providing interactivity that exceeds that of comparable software designed for VFX usage.
@@
 I wanted to include a node-graph style interface within my application, as these are commonly used within existing artist tools (Autodesk Maya, Unreal Engine), and would encourage modularity within my API.
-Qt does not natively provide this kind of interface, so I made use of NodeEditor~\cite{Pinaev2017}, an existing Qt-based library that provides this functionality.
+Qt does not natively provide this kind of interface, so I made use of \textit{NodeEditor}~\cite{Pinaev2017}, an existing Qt-based library that provides this functionality.
 
@@
 Another desirable addition to my system would be viewport manipulators, similar to the 'Groomable Splines' feature in XGen.
-This could allow me to implement techniques such as those described in Skippy~\cite{skippy}, which would provide my system with a unique user experience compared to commercially available systems.
+This could allow me to implement techniques such as those described in \textit{Skippy}~\cite{skippy}, which would provide my system with a unique user experience compared to commercially available systems.
 

### 2018-03-19 docs/paper.tex
docs tweaks

@@
 After I had implemented some of the features Alin requested, I decided that the system was in a satisfactory state for the purposes of this assignment.
-The following Figures~\ref{figure:screenshots} and ~\ref{figure:screenshot_gui} show examples of the kind of visual output my system is capable of producing, and the accompanying node graph setup used to create them.
+The following Figures~\ref{figure:screenshots} and~\ref{figure:screenshot_gui} show examples of the kind of visual output my system is capable of producing, and the accompanying node graph setup used to create them.
 

### 2018-03-19 include/Renderable.h
removed prints

@@ -17,7 +17,7 @@ class Renderable
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief constructor
 		//----------------------------------------------------------------------------------------------------------------------
-		Renderable(){std::cout<<"Renderable()\n";}
+		Renderable(){}//std::cout<<"Renderable()\n";}
 		//----------------------------------------------------------------------------------------------------------------------
 		/// @brief destructor
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-19 src/BendCurveOperatorModel.cpp

@@ -38,7 +38,7 @@ void BendCurveOperatorModel::operateCurves()
 			break;
 
 		case SSBO:
-			std::cout<<"processing bend curves SSBO\n";
+			// std::cout<<"processing bend curves SSBO\n";
 			ComputeShaderManager::getInstance()->bendCurvesOperator(m_curvesSSBO, bendDirection, intensity);
 			break;
 	}

### 2018-03-19 src/ClumpCurveOperatorModel.cpp

@@ -124,7 +124,7 @@ void ClumpCurveOperatorModel::operateCurves()
 
 	if (curveType == SSBO && clumpCurveType == SSBO)
 	{
-		std::cout<<"processing clump curves SSBO: "<<m_curvesSSBO<<" & "<<m_clumpCurvesSSBO<<"\n";
+		// std::cout<<"processing clump curves SSBO: "<<m_curvesSSBO<<" & "<<m_clumpCurvesSSBO<<"\n";
 		ComputeShaderManager::getInstance()->clumpCurvesOperator(m_curvesSSBO, m_clumpCurvesSSBO, m_ui->envelopeSpinBox->value(), m_ui->preserveLengthCheckBox->isChecked());
 	}
 	else

### 2018-03-19 src/ComputeShaderManager.cpp

@@ -53,7 +53,7 @@ void ComputeShaderManager::recompileShaders()
 
 void ComputeShaderManager::createCurvesSSBO(unsigned int &buffer, Curves _curves)
 {
-	std::cout<<"creating an SSBO for curves\n";
+	// std::cout<<"creating an SSBO for curves\n";
 
 	if (buffer)
 		glDeleteBuffers(1, &buffer);
@@ -80,7 +80,7 @@ void ComputeShaderManager::createCurvesSSBO(unsigned int &buffer, Curves _curves
 
 void ComputeShaderManager::createMeshSSBO(unsigned int &buffer, Mesh &mesh)
 {
-	std::cout<<"creating an SSBO for mesh\n";
+	// std::cout<<"creating an SSBO for mesh\n";
 
 	if (buffer)
 		glDeleteBuffers(1, &buffer);
@@ -123,14 +123,14 @@ void ComputeShaderManager::createMeshSSBO(unsigned int &buffer, Mesh &mesh)
 
 void ComputeShaderManager::copyCurvesSSBO(unsigned int src, unsigned int &dst)
 {
-	std::cout<<"copying an SSBO into another: "<<src<<" to "<<dst<<"\n";
+	// std::cout<<"copying an SSBO into another: "<<src<<" to "<<dst<<"\n";
 	// get size of src buffer
 	GLint srcBufferSize = 0;
 	glBindBuffer(GL_SHADER_STORAGE_BUFFER, src);
 	glGetBufferParameteriv(GL_SHADER_STORAGE_BUFFER, GL_BUFFER_SIZE, &srcBufferSize);
 	glBindBuffer(GL_SHADER_STORAGE_BUFFER, 0);
 
-	std::cout<<"src buffer size : "<<srcBufferSize<<"\n";
+	// std::cout<<"src buffer size : "<<srcBufferSize<<"\n";
 
 	// first check if dst is a buffer and if they are equal sized
 	// if dst buffer exist check if buffers are equal size
@@ -170,8 +170,8 @@ void ComputeShaderManager::copyCurvesSSBO(unsigned int src, unsigned int &dst)
 	GLint newBufferSize = 0;
 	glBindBuffer(GL_SHADER_STORAGE_BUFFER, dst);
 	glGetBufferParameteriv(GL_SHADER_STORAGE_BUFFER, GL_BUFFER_SIZE, &newBufferSize);
[…]

### 2018-03-19 src/CurvesRenderable.cpp

@@ -20,7 +20,7 @@ CurvesRenderable::CurvesRenderable()
 
 CurvesRenderable::~CurvesRenderable()
 {
-	std::cout<<"~CurvesRenderable "<<this<<"\n";
+	// std::cout<<"~CurvesRenderable "<<this<<"\n";
 }
 
 //----------------------------------------------------------------------------------------------------------------------
@@ -41,9 +41,9 @@ void CurvesRenderable::generate()
 			glBindBuffer(GL_SHADER_STORAGE_BUFFER, m_SSBO);
 			glGetBufferParameteriv(GL_SHADER_STORAGE_BUFFER, GL_BUFFER_SIZE, &SSBOSize);
 			glBindBuffer(GL_SHADER_STORAGE_BUFFER, 0);
-			std::cout<<"SSBO size : "<<SSBOSize<<"\n";
+			// std::cout<<"SSBO size : "<<SSBOSize<<"\n";
 			m_indices = SSBOSize / (sizeof(glm::vec4) * 6);
-			std::cout<<"m_indices from SSBO size : "<<m_indices<<"\n";
+			// std::cout<<"m_indices from SSBO size : "<<m_indices<<"\n";
 			break;
 	}
 
@@ -64,7 +64,7 @@ void CurvesRenderable::cleanUp()
 {
 	if (isConstructed)
 	{
-		std::cout<<"deleting SSBO: "<<m_SSBO<<" and empty VAO: "<<m_emptyVAO<<"\n";
+		// std::cout<<"deleting SSBO: "<<m_SSBO<<" and empty VAO: "<<m_emptyVAO<<"\n";
 
 		if (m_SSBO)
 			glDeleteBuffers(1, &m_SSBO);
@@ -79,7 +79,7 @@ void CurvesRenderable::cleanUp()
 
 void CurvesRenderable::draw()
 {
-	std::cout<<"drawing "<<this<<" & "<<m_indices<<" indices of "<<m_SSBO<<"\n";
+	// std::cout<<"drawing "<<this<<" & "<<m_indices<<" indices of "<<m_SSBO<<"\n";
 
 	glBindBufferBase(GL_SHADER_STORAGE_BUFFER, 0, m_SSBO);
 

### 2018-03-19 src/CurvesRendererModel.cpp

@@ -27,7 +27,7 @@ CurvesRendererModel::CurvesRendererModel() :
 
 CurvesRendererModel::~CurvesRendererModel()
 {
-	std::cout<<"~CurvesRendererModel, setting "<<m_curvesRenderable<<" to be deleted\n";
+	// std::cout<<"~CurvesRendererModel, setting "<<m_curvesRenderable<<" to be deleted\n";
 	m_curvesRenderable->setToBeDeleted();
 	delete m_ui;
 	// delete m_embedded; // for some reason these cause crash

### 2018-03-19 src/NoiseCurveOperatorModel.cpp

@@ -30,11 +30,11 @@ void NoiseCurveOperatorModel::operateCurves()
 	switch (std::static_pointer_cast<CurvesData>(_nodeData)->curveType())
 	{
 		case CPU:
-			std::cout<<"processing noise curves not possible on CPU\n";
+			// std::cout<<"processing noise curves not possible on CPU\n";
 			break;
 
 		case SSBO:
-			std::cout<<"processing noise curves SSBO\n";
+			// std::cout<<"processing noise curves SSBO\n";
 			ComputeShaderManager::getInstance()->noiseCurvesOperator(m_curvesSSBO, intensity);
 			break;
 	}

### 2018-03-19 src/ObjMeshModel.cpp

@@ -60,7 +60,7 @@ bool ObjMeshModel::eventFilter(QObject *object, QEvent *event)
 				QDir::currentPath() + "/meshes",
 				tr("Mesh Files (*.obj)"));
 
-			ComputeShaderManager::getInstance()->startTimer();
+			// ComputeShaderManager::getInstance()->startTimer();
 
 			m_label->setText(fileName);
 

### 2018-03-19 src/RenderableManager.cpp

@@ -69,7 +69,7 @@ void RenderableManager::registerRenderable(std::shared_ptr<Renderable> _renderab
 void RenderableManager::deregisterRenderable(std::shared_ptr<Renderable> _renderable)
 {
 	m_renderables.erase(std::remove(m_renderables.begin(), m_renderables.end(), _renderable), m_renderables.end());
-	std::cout<<"deregistered "<<_renderable<<"\n";
+	// std::cout<<"deregistered "<<_renderable<<"\n";
 }
 
 //----------------------------------------------------------------------------------------------------------------------

### 2018-03-19 include/BendCurveOperatorModel.h
fixed saving and loading

@@ -31,6 +31,10 @@ class BendCurveOperatorModel : public AbstractCurveOperatorModel
 		//----------------------------------------------------------------------------------------------------------------------
 		QWidget* embeddedWidget() override {return m_spinbox;};
 		//----------------------------------------------------------------------------------------------------------------------
+		QJsonObject save() const override;
+		//----------------------------------------------------------------------------------------------------------------------
+		void restore(QJsonObject const &p) override;
+		//----------------------------------------------------------------------------------------------------------------------
 
 	private:
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-19 include/ClumpCurveOperatorModel.h

@@ -41,6 +41,10 @@ class ClumpCurveOperatorModel : public AbstractCurveOperatorModel
 		//----------------------------------------------------------------------------------------------------------------------
 		virtual QString portCaption(PortType portType, PortIndex portIndex) const override;
 		//----------------------------------------------------------------------------------------------------------------------
+		QJsonObject save() const override;
+		//----------------------------------------------------------------------------------------------------------------------
+		void restore(QJsonObject const &p) override;
+		//----------------------------------------------------------------------------------------------------------------------
 
 	private:
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-19 include/ComputeShaderManager.h

@@ -70,14 +70,14 @@ class ComputeShaderManager
 		void clumpCurvesOperator(unsigned int curvesSSBO, unsigned int clumpCurvesSSBO, float envelope, bool preserveLength);
 
 
-		//----------------------------------------------------------------------------------------------------------------------
-		/// @brief start chrono timer
-		//----------------------------------------------------------------------------------------------------------------------
-		void startTimer();
-		//----------------------------------------------------------------------------------------------------------------------
-		/// @brief end timer and print elapsed
-		//----------------------------------------------------------------------------------------------------------------------
-		void endTimer();
+		// //----------------------------------------------------------------------------------------------------------------------
+		// /// @brief start chrono timer
+		// //----------------------------------------------------------------------------------------------------------------------
+		// void startTimer();
+		// //----------------------------------------------------------------------------------------------------------------------
+		// /// @brief end timer and print elapsed
+		// //----------------------------------------------------------------------------------------------------------------------
+		// void endTimer();
 
 	private:
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-19 include/CurvesRendererModel.h

@@ -55,6 +55,11 @@ class CurvesRendererModel : public NodeDataModel
 		QWidget* embeddedWidget() override {return m_embedded;};
 		//----------------------------------------------------------------------------------------------------------------------
 		bool resizable() const override {return false;};
+		//----------------------------------------------------------------------------------------------------------------------
+		QJsonObject save() const override;
+		//----------------------------------------------------------------------------------------------------------------------
+		void restore(QJsonObject const &p) override;
+		//----------------------------------------------------------------------------------------------------------------------
 
 	public slots:
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-19 include/MeshRendererModel.h

@@ -55,6 +55,11 @@ class MeshRendererModel : public NodeDataModel
 		QWidget* embeddedWidget() override {return m_embedded;};
 		//----------------------------------------------------------------------------------------------------------------------
 		bool resizable() const override {return false;};
+		//----------------------------------------------------------------------------------------------------------------------
+		QJsonObject save() const override;
+		//----------------------------------------------------------------------------------------------------------------------
+		void restore(QJsonObject const &p) override;
+		//----------------------------------------------------------------------------------------------------------------------
 
 	public slots:
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-19 include/NoiseCurveOperatorModel.h

@@ -31,6 +31,10 @@ class NoiseCurveOperatorModel : public AbstractCurveOperatorModel
 		//----------------------------------------------------------------------------------------------------------------------
 		QWidget* embeddedWidget() override {return m_spinbox;};
 		//----------------------------------------------------------------------------------------------------------------------
+		QJsonObject save() const override;
+		//----------------------------------------------------------------------------------------------------------------------
+		void restore(QJsonObject const &p) override;
+		//----------------------------------------------------------------------------------------------------------------------
 
 	private:
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-19 include/RandomDistributorModel.h

@@ -33,6 +33,10 @@ class RandomDistributorModel : public AbstractDistributorModel
 		//----------------------------------------------------------------------------------------------------------------------
 		std::shared_ptr<NodeData> outData(PortIndex port) override;
 		//----------------------------------------------------------------------------------------------------------------------
+		QJsonObject save() const override;
+		//----------------------------------------------------------------------------------------------------------------------
+		void restore(QJsonObject const &p) override;
+		//----------------------------------------------------------------------------------------------------------------------
 
 	protected:
 		//----------------------------------------------------------------------------------------------------------------------

### 2018-03-19 nodeeditor/FlowScene.cpp

@@ -410,7 +410,7 @@ save() const
   QString fileName =
     QFileDialog::getSaveFileName(nullptr,
                                  tr("Open Flow Scene"),
-                                 QDir::homePath(),
+                                 QDir::currentPath() + "/tests",
                                  tr("Flow Scene Files (*.flow)"));
 
   if (!fileName.isEmpty())
@@ -438,7 +438,7 @@ load()
   QString fileName =
     QFileDialog::getOpenFileName(nullptr,
                                  tr("Open Flow Scene"),
-                                 QDir::homePath(),
+                                 QDir::currentPath() + "/tests",
                                  tr("Flow Scene Files (*.flow)"));
 
   if (!QFileInfo::exists(fileName))

### 2018-03-19 src/AbstractCurveOperatorModel.cpp

@@ -112,15 +112,18 @@ void AbstractCurveOperatorModel::operateCurves()
 
 void AbstractCurveOperatorModel::resetCurves()
 {
-	switch (_nodeData->curveType())
+	if(_nodeData)
 	{
-		case CPU:
-			m_curves = _nodeData->curves();
-			break;
+		switch (_nodeData->curveType())
+		{
+			case CPU:
+				m_curves = _nodeData->curves();
+				break;
 
-		case SSBO:
-			ComputeShaderManager::getInstance()->copyCurvesSSBO(_nodeData->curvesSSBO(), m_curvesSSBO);
-			break;
+			case SSBO:
+				ComputeShaderManager::getInstance()->copyCurvesSSBO(_nodeData->curvesSSBO(), m_curvesSSBO);
+				break;
+		}
 	}
 }
 

### 2018-03-19 src/BendCurveOperatorModel.cpp

@@ -20,27 +20,46 @@ BendCurveOperatorModel::~BendCurveOperatorModel()
 	// delete m_spinbox;
 }
 
+QJsonObject BendCurveOperatorModel::save() const
+{
+	QJsonObject modelJson = NodeDataModel::save();
+
+	modelJson["value"] = m_spinbox->value();
+
+	return modelJson;
+}
+
+void BendCurveOperatorModel::restore(QJsonObject const &p)
+{
+	QJsonValue r = p["value"];
+	if (!r.isUndefined())
+		m_spinbox->setValue(r.toDouble());
+
+}
+
 void BendCurveOperatorModel::operateCurves()
 {
-	float intensity = m_spinbox->value();
-	glm::vec3 bendDirection = glm::vec3(0.0, -1.0, 0.0);
-
-	switch (std::static_pointer_cast<CurvesData>(_nodeData)->curveType())
+	if(_nodeData)
 	{
-		case CPU:
-			for (auto &curve : m_curves.m_curves)
-			{
-				for (unsigned int j = 1; j < 5; ++j)
+		float intensity = m_spinbox->value();
+		glm::vec3 bendDirection = glm::vec3(0.0, -1.0, 0.0);
+
+		switch (std::static_pointer_cast<CurvesData>(_nodeData)->curveType())
+		{
+			case CPU:
+				for (auto &curve : m_curves.m_curves)
 				{
-					curve.vertices[j] += float(j) * float(j) * intensity * bendDirection;
+					for (unsigned int j = 1; j < 5; ++j)
+					{
+						curve.vertices[j] += float(j) * float(j) * intensity * bendDirection;
+					}
 				}
-			}
-			break;
+				break;
 
-		case SSBO:
-			// std::cout<<"processing bend curves SSBO\n";
-			ComputeShaderManager::getInstance()->bendCurvesOperator(m_curvesSSBO, bendDirection, intensity);
-			break;
+			case SSBO:
+				// std::cout<<"processing bend curves SSBO\n";
+				ComputeShaderManager::getInstance()->bendCurvesOperator(m_curvesSSBO, bendDirection, intensity);
+				break;
+		}
 	}
-
 }

### 2018-03-19 src/ClumpCurveOperatorModel.cpp

@@ -28,6 +28,27 @@ ClumpCurveOperatorModel::~ClumpCurveOperatorModel()
 	delete m_ui;
 }
 
+QJsonObject ClumpCurveOperatorModel::save() const
+{
+	QJsonObject modelJson = NodeDataModel::save();
+
+	modelJson["envelope"] = m_ui->envelopeSpinBox->value();
+	modelJson["preserve"] = m_ui->preserveLengthCheckBox->isChecked();
+
+	return modelJson;
+}
+
+void ClumpCurveOperatorModel::restore(QJsonObject const &p)
+{
+	QJsonValue r = p["envelope"];
+	if (!r.isUndefined())
+		m_ui->envelopeSpinBox->setValue(r.toDouble());
+
+	QJsonValue v = p["preserve"];
+	if (!v.isUndefined())
+		m_ui->preserveLengthCheckBox->setChecked(v.toBool());
+}
+
 QString ClumpCurveOperatorModel::portCaption(PortType portType, PortIndex portIndex) const
 {
 	switch (portType)
@@ -103,32 +124,38 @@ void ClumpCurveOperatorModel::setInData(std::shared_ptr<NodeData> nodeData, Port
 
 void ClumpCurveOperatorModel::resetCurves()
 {
-	AbstractCurveOperatorModel::resetCurves();
-
-	switch (m_clumpNodeData->curveType())
+	if(_nodeData)
 	{
-		case CPU:
-			m_clumpCurves = m_clumpNodeData->curves();
-			break;
+		AbstractCurveOperatorModel::resetCurves();
 
-		case SSBO:
-			ComputeShaderManager::getInstance()->copyCurvesSSBO(m_clumpNodeData->curvesSSBO(), m_clumpCurvesSSBO);
-			break;
+		switch (m_clumpNodeData->curveType())
+		{
+			case CPU:
+				m_clumpCurves = m_clumpNodeData->curves();
+				break;
+
+			case SSBO:
+				ComputeShaderManager::getInstance()->copyCurvesSSBO(m_clumpNodeData->curvesSSBO(), m_clumpCurvesSSBO);
+				break;
+		}
 	}
 }
 
 void ClumpCurveOperatorModel::operateCurves()
 {
-	auto curveType = std::static_pointer_cast<CurvesData>(_nodeData)->curveType();
-	auto clumpCurveType = std::static_pointer_cast<CurvesData>(m_clumpNodeData)->curveType();
[…]

### 2018-03-19 src/ComputeShaderManager.cpp

@@ -276,16 +276,16 @@ void ComputeShaderManager::clumpCurvesOperator(unsigned int curvesSSBO, unsigned
 
 //----------------------------------------------------------------------------------------------------------------------
 
-void ComputeShaderManager::startTimer()
-{
-	m_timer = std::chrono::system_clock::now();
-}
+// void ComputeShaderManager::startTimer()
+// {
+// 	m_timer = std::chrono::system_clock::now();
+// }
 
-//----------------------------------------------------------------------------------------------------------------------
+// //----------------------------------------------------------------------------------------------------------------------
 
-void ComputeShaderManager::endTimer()
-{
-	auto endTimer = std::chrono::system_clock::now();
-	auto elapsed = std::chrono::duration_cast<std::chrono::milliseconds>(endTimer - m_timer);
-	std::cout << "timer measured " << elapsed.count() << "ms\n";
-}
+// void ComputeShaderManager::endTimer()
+// {
+// 	auto endTimer = std::chrono::system_clock::now();
+// 	auto elapsed = std::chrono::duration_cast<std::chrono::milliseconds>(endTimer - m_timer);
+// 	std::cout << "timer measured " << elapsed.count() << "ms\n";
+// }

### 2018-03-19 src/CurvesRendererModel.cpp

@@ -34,6 +34,57 @@ CurvesRendererModel::~CurvesRendererModel()
 	// m_embedded->deleteLater();
 }
 
+QJsonObject CurvesRendererModel::save() const
+{
+	QJsonObject modelJson = NodeDataModel::save();
+
+	modelJson["red"] = m_ui->redAmountDoubleSpinBox->value();
+	modelJson["green"] = m_ui->greenAmountDoubleSpinBox->value();
+	modelJson["blue"] = m_ui->blueAmountDoubleSpinBox->value();
+	modelJson["width"] = m_ui->widthDoubleSpinBox->value();
+	modelJson["roughness"] = m_ui->roughnessDoubleSpinBox->value();
+	modelJson["metallic"] = m_ui->metallicDoubleSpinBox->value();
+	modelJson["ribbons"] = m_ui->ribbonsCheckBox->isChecked();
+	modelJson["visibiliy"] = m_ui->visibilityCheckBox->isChecked();
+
+	return modelJson;
+}
+
+void CurvesRendererModel::restore(QJsonObject const &p)
+{
+	QJsonValue r = p["red"];
+	if (!r.isUndefined())
+		m_ui->redAmountDoubleSpinBox->setValue(r.toDouble());
+
+	QJsonValue g = p["green"];
+	if (!g.isUndefined())
+		m_ui->greenAmountDoubleSpinBox->setValue(g.toDouble());
+
+	QJsonValue b = p["blue"];
+	if (!b.isUndefined())
+		m_ui->blueAmountDoubleSpinBox->setValue(b.toDouble());
+
+	QJsonValue w = p["width"];
+	if (!w.isUndefined())
+		m_ui->widthDoubleSpinBox->setValue(w.toDouble());
+
+	QJsonValue ro = p["roughness"];
+	if (!ro.isUndefined())
+		m_ui->roughnessDoubleSpinBox->setValue(ro.toDouble());
+
+	QJsonValue m = p["metallic"];
+	if (!m.isUndefined())
+		m_ui->metallicDoubleSpinBox->setValue(m.toDouble());
+
+	QJsonValue ri = p["ribbons"];
+	if (!ri.isUndefined())
+		m_ui->ribbonsCheckBox->setChecked(ri.toBool());
+
+	QJsonValue v = p["visibiliy"];
+	if (!v.isUndefined())
+		m_ui->visibilityCheckBox->setChecked(v.toBool());
+}
+
[…]

### 2018-03-19 src/MainWindow.cpp

@@ -23,7 +23,7 @@ static std::shared_ptr<DataModelRegistry> registerDataModels()
 	ret->registerModel<ObjMeshModel>("Mesh Loaders");
 	ret->registerModel<MeshRendererModel>("Renderers");
 	ret->registerModel<CurvesRendererModel>("Renderers");
-	ret->registerModel<BasicDistributorModel>("Distributors");
+	// ret->registerModel<BasicDistributorModel>("Distributors");
 	ret->registerModel<RandomDistributorModel>("Distributors");
 	ret->registerModel<BendCurveOperatorModel>("Curve Operators");
 	ret->registerModel<NoiseCurveOperatorModel>("Curve Operators");

### 2018-03-19 src/MeshRendererModel.cpp

@@ -32,6 +32,47 @@ MeshRendererModel::~MeshRendererModel()
 	// m_embedded->deleteLater();
 }
 
+QJsonObject MeshRendererModel::save() const
+{
+	QJsonObject modelJson = NodeDataModel::save();
+
+	modelJson["red"] = m_ui->redAmountDoubleSpinBox->value();
+	modelJson["green"] = m_ui->greenAmountDoubleSpinBox->value();
+	modelJson["blue"] = m_ui->blueAmountDoubleSpinBox->value();
+	modelJson["roughness"] = m_ui->roughnessDoubleSpinBox->value();
+	modelJson["metallic"] = m_ui->metallicDoubleSpinBox->value();
+	modelJson["visibiliy"] = m_ui->visibilityCheckBox->isChecked();
+
+	return modelJson;
+}
+
+void MeshRendererModel::restore(QJsonObject const &p)
+{
+	QJsonValue r = p["red"];
+	if (!r.isUndefined())
+		m_ui->redAmountDoubleSpinBox->setValue(r.toDouble());
+
+	QJsonValue g = p["green"];
+	if (!g.isUndefined())
+		m_ui->greenAmountDoubleSpinBox->setValue(g.toDouble());
+
+	QJsonValue b = p["blue"];
+	if (!b.isUndefined())
+		m_ui->blueAmountDoubleSpinBox->setValue(b.toDouble());
+
+	QJsonValue ro = p["roughness"];
+	if (!ro.isUndefined())
+		m_ui->roughnessDoubleSpinBox->setValue(ro.toDouble());
+
+	QJsonValue m = p["metallic"];
+	if (!m.isUndefined())
+		m_ui->metallicDoubleSpinBox->setValue(m.toDouble());
+
+	QJsonValue v = p["visibiliy"];
+	if (!v.isUndefined())
+		m_ui->visibilityCheckBox->setChecked(v.toBool());
+}
+
 unsigned int MeshRendererModel::nPorts(PortType portType) const
 {
 	unsigned int result = 1;

### 2018-03-19 src/NoiseCurveOperatorModel.cpp

@@ -20,23 +20,41 @@ NoiseCurveOperatorModel::~NoiseCurveOperatorModel()
 	// delete m_spinbox;
 }
 
+QJsonObject NoiseCurveOperatorModel::save() const
+{
+	QJsonObject modelJson = NodeDataModel::save();
+
+	modelJson["value"] = m_spinbox->value();
+
+	return modelJson;
+}
+
+void NoiseCurveOperatorModel::restore(QJsonObject const &p)
+{
+	QJsonValue r = p["value"];
+	if (!r.isUndefined())
+		m_spinbox->setValue(r.toDouble());
+}
+
 void NoiseCurveOperatorModel::operateCurves()
 {
-	float intensity = m_spinbox->value();
-
-	if (intensity == 0.0)
-		return;
-
-	switch (std::static_pointer_cast<CurvesData>(_nodeData)->curveType())
+	if(_nodeData)
 	{
-		case CPU:
-			// std::cout<<"processing noise curves not possible on CPU\n";
-			break;
+		float intensity = m_spinbox->value();
 
-		case SSBO:
-			// std::cout<<"processing noise curves SSBO\n";
-			ComputeShaderManager::getInstance()->noiseCurvesOperator(m_curvesSSBO, intensity);
-			break;
+		if (intensity == 0.0)
+			return;
+
+		switch (std::static_pointer_cast<CurvesData>(_nodeData)->curveType())
+		{
+			case CPU:
+				// std::cout<<"processing noise curves not possible on CPU\n";
+				break;
+
+			case SSBO:
+				// std::cout<<"processing noise curves SSBO\n";
+				ComputeShaderManager::getInstance()->noiseCurvesOperator(m_curvesSSBO, intensity);
+				break;
+		}
 	}
-
 }

### 2018-03-19 src/RandomDistributorModel.cpp

@@ -37,6 +37,37 @@ RandomDistributorModel::~RandomDistributorModel()
 	// m_embedded->deleteLater();
 }
 
+QJsonObject RandomDistributorModel::save() const
+{
+	QJsonObject modelJson = NodeDataModel::save();
+
+	modelJson["count"] = m_ui->countSpinBox->value();
+	modelJson["mode"] = m_ui->modeCheckBox->isChecked();
+	modelJson["length"] = m_ui->lengthSpinBox->value();
+	modelJson["variation"] = m_ui->variationSpinBox->value();
+
+	return modelJson;
+}
+
+void RandomDistributorModel::restore(QJsonObject const &p)
+{
+	QJsonValue c = p["count"];
+	if (!c.isUndefined())
+		m_ui->countSpinBox->setValue(c.toInt());
+
+	QJsonValue m = p["mode"];
+	if (!m.isUndefined())
+		m_ui->modeCheckBox->setChecked(m.toBool());
+
+	QJsonValue l = p["length"];
+	if (!l.isUndefined())
+		m_ui->lengthSpinBox->setValue(l.toDouble());
+
+	QJsonValue v = p["variation"];
+	if (!v.isUndefined())
+		m_ui->variationSpinBox->setValue(v.toDouble());
+}
+
 void RandomDistributorModel::meshChanged()
 {
 	ComputeShaderManager::getInstance()->createMeshSSBO(m_meshSSBOID, m_mesh);

### 2018-03-19 src/Viewport.cpp

@@ -114,7 +114,7 @@ void Viewport::paintGL()
 	loadMatricesToRenderManager();
 	RenderableManager::getInstance()->drawRenderables();
 
-	ComputeShaderManager::getInstance()->endTimer();
+	// ComputeShaderManager::getInstance()->endTimer();
 }
 
 //----------------------------------------------------------------------------------------------------------------------
